---
title: "Analysis Results"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

# Setup

## -- Load Analyses

```{r}
#setwd("~/leukem_ai/R")
#source("inner_cv_analysis.R")
#inner_cv_results_merged <- readRDS("../data/out/inner_cv/inner_cv_results_10feb2026_merged_maxprob.rds")
#inner_cv_results_merged_summed <- readRDS("../data/out/inner_cv/inner_cv_results_10feb2026_merged_summed.rds")

#inner_cv_results_merged$final_results
#inner_cv_results_merged_summed$final_results
```



```{r}
#source("outer_cv_analysis.R")
#outer_cv_results_unmerged[["performance_summaries"]]
#outer_cv_results_merged[["performance_summaries"]]
```

```{r}
outer_unmerged <-  readRDS("../data/out/outer_cv/outer_cv_results_10feb26_unmerged_maxprob.rds")
outer_merged <- readRDS("../data/out/outer_cv/outer_cv_results_10feb26_merged_summed.rds")
```



```{r}
#source("train_test_analysis.R")
```

# Load data
```{r}
modify_classes <- function(vector) {
    # Convert factor to character to avoid invalid factor level warnings
    was_factor <- is.factor(vector)
    vector <- as.character(vector)
    
    vector[grepl("MDS|TP53", vector)] <- "MDS.r"
    vector[!grepl("MLLT3", vector) & grepl("KMT2A", vector)] <- "other.KMT2A"
    vector[grepl("MECOM", vector)] <- "MECOM"

    
    # Optionally convert back to factor if input was factor
    if (was_factor) {
      vector <- factor(vector)
    }
    vector
  }

# Filters
DATA_FILTERS <- list(
  min_samples_per_subtype = 10,
  excluded_subtypes = c("AML NOS", "Missing data", "Multi"),
  selected_studies = c(
    "TCGA-LAML",
    "LEUCEGENE",
    "BEATAML1.0-COHORT",
    "AAML0531",
    "AAML1031",
    "AAML03P1",
    "100LUMC"
  )
)

# Load mapping of class labels to numeric labels
label_mapping <- read.csv("../data/label_mapping_all.csv")

# Load leukemia subtype data
leukemia_subtypes <- read.csv("../data/rgas_10feb26.csv")$ICC_Subtype

# Load study metadata
meta <- read.csv("../data/meta_20aug25.csv")
study_names <- meta$Studies

# Filter data based on criteria
subtypes_with_sufficient_samples <- names(which(table(leukemia_subtypes) >= DATA_FILTERS$min_samples_per_subtype))
filter <- which(
  leukemia_subtypes %in% subtypes_with_sufficient_samples &
    !leukemia_subtypes %in% DATA_FILTERS$excluded_subtypes &
    study_names %in% DATA_FILTERS$selected_studies
)

leukemia_subtypes_mod <- modify_classes(leukemia_subtypes)

filtered_leukemia_subtypes <- leukemia_subtypes[filter]
filtered_leukemia_subtypes[filtered_leukemia_subtypes == "MECOM fusion"] <- "MECOM rearrangement"
# Load study metadata
filtered_study_names <- study_names[filter]

meta <- meta[filter, ]
meta$classes <- filtered_leukemia_subtypes


```

```{r}
dir.create("../writing/tables_new")
dir.create("../writing/figures_new")
```

# Training Cohort Tables
```{r}
length(unique(filtered_leukemia_subtypes))

table(filtered_leukemia_subtypes, meta$Studies)
```

## -- Supplementary Table 1: All Subtypes (Included and Excluded)

```{r results='asis'}
library(dplyr)
library(knitr)
library(kableExtra)

# Create data frame for included classes
included_df <- data.frame(
  Subtype = filtered_leukemia_subtypes,
  Cohort = filtered_study_names,
  stringsAsFactors = FALSE
)

# Create data frame for excluded classes
excluded_df <- data.frame(
  Subtype = leukemia_subtypes[-filter],
  Cohort = study_names[-filter],
  stringsAsFactors = FALSE
)

# Define cohort order (adults first, then pediatric)
cohort_order <- c(
  # Adult cohorts
  "TCGA-LAML",
  "LEUCEGENE",
  "BEATAML1.0-COHORT",
  "100LUMC",
  # Pediatric cohorts
  "AAML0531",
  "AAML1031",
  "AAML03P1"
)

# Function to create cross-tabulation table
create_subtype_table <- function(df, cohort_order) {
  # Create cross-tabulation
  ct <- table(df$Subtype, df$Cohort)
  ct_df <- as.data.frame.matrix(ct)
  
  # Ensure all cohorts are present (even if 0 counts)
  for (cohort in cohort_order) {
    if (!cohort %in% colnames(ct_df)) {
      ct_df[[cohort]] <- 0
    }
  }
  
  # Reorder columns according to cohort_order
  ct_df <- ct_df[, cohort_order, drop = FALSE]
  
  # Add Total column
  ct_df$Total <- rowSums(ct_df)
  
  # Sort rows by Total (high to low)
  ct_df <- ct_df[order(-ct_df$Total), , drop = FALSE]
  
  # Move rownames to a column
  ct_df <- cbind(Subtype = rownames(ct_df), ct_df)
  rownames(ct_df) <- NULL
  
  return(ct_df)
}

# Create tables for included and excluded classes

# For included, we need to show merged classes with breakdowns and
# also group rare recurrent translocations into a single AML subgroup.
included_df_modified <- included_df
included_df_modified$Subtype_Original <- included_df$Subtype

# Start from the standard merged labels (MDS-related, other KMT2A, MECOM)
included_df_modified$Subtype_Merged <- modify_classes(included_df$Subtype)

# Define rare recurrent translocation subtypes to be grouped.
# Allow for both "::" and ".." / comma variants so this works with
# whichever naming is present in filtered_leukemia_subtypes.
rare_tx_subtypes <- c(
  "CBFA2T3::GLIS2",        "CBFA2T3..GLIS2",
  "NUP98::KDM5A",          "NUP98..KDM5A",
  "KAT6A::CREBBP",         "KAT6A..CREBBP",
  "ETV6::MNX1",            "ETV6..MNX1",
  "FUS::ERG", "RBM15::MRTF1", "NUP98::NSD1",
  "NUP98, other partners", "NUP98..other.partners", "PICALM::MLLT10", "PICALM..MLLT10"
)

# Map these to a single merged group
included_df_modified$Subtype_Merged[
  included_df_modified$Subtype_Original %in% rare_tx_subtypes
] <- "other.rare.tx"

# Create table with merged classes
included_table_merged <- create_subtype_table(
  data.frame(Subtype = included_df_modified$Subtype_Merged, 
             Cohort = included_df_modified$Cohort), 
  cohort_order
)

# Rename merged subtypes for display
included_table_merged$Subtype <- gsub("MDS\\.r", "MDS-related", included_table_merged$Subtype)
included_table_merged$Subtype <- gsub("other\\.KMT2A", "Other KMT2A rearrangements", included_table_merged$Subtype)
included_table_merged$Subtype[included_table_merged$Subtype == "MECOM"] <- "MECOM rearrangement"
included_table_merged$Subtype <- gsub(
  "other\\.rare\\.tx",
  "AML with other rare recurring translocations",
  included_table_merged$Subtype
)

# Custom display order for included AML subtypes
# (controls order in both Supplementary Table 1 and Table 1)
# Use the actual row labels from the table.
desired_included_order <- c(
  "APL, t(15;17)/PML::RARA",                 # Acute promyelocytic leukemia
  "AML with t(8;21)/RUNX1::RUNX1T1",
  "AML with inv(16)/t(16;16)/CBFB::MYH11",
  "AML with t(9;11)/MLLT3::KMT2A",
  "Other KMT2A rearrangements",
  "AML with t(6;9)/DEK::NUP214",
  # MECOM rearrangements grouped at merged level
  "MECOM rearrangement",
  "GATA2;MECOM",
  "MECOM, other",
  "AML with other rare recurring translocations",
  "AML with t(9;22)/BCR::ABL1",             # May be absent; ordering still safe
  "NUP98::NSD1",
  "AML with mutated NPM1",
  "AML with in-frame bZIP CEBPA",
  "MDS-related"                             # MDS/AML-related entities at the end
)

included_table_merged$Subtype <- factor(
  included_table_merged$Subtype,
  levels = c(
    desired_included_order,
    setdiff(included_table_merged$Subtype, desired_included_order)
  )
)
included_table_merged <- included_table_merged[order(included_table_merged$Subtype), ]
included_table_merged$Subtype <- as.character(included_table_merged$Subtype)

# Create breakdown tables for merged classes
# MDS.r/MECOM breakdown
mds_subtypes <- included_df_modified$Subtype_Original[included_df_modified$Subtype_Merged == "MDS.r"]
if (length(mds_subtypes) > 0) {
  mds_breakdown <- create_subtype_table(
    data.frame(Subtype = mds_subtypes, 
               Cohort = included_df_modified$Cohort[included_df_modified$Subtype_Merged == "MDS.r"]),
    cohort_order
  )
  mds_breakdown$Subtype <- paste0("<span style='padding-left:20px;'>└─ ", mds_breakdown$Subtype, "</span>")
}

# other.KMT2A breakdown
other_kmt2a_subtypes <- included_df_modified$Subtype_Original[included_df_modified$Subtype_Merged == "other.KMT2A"]
if (length(other_kmt2a_subtypes) > 0) {
  other_kmt2a_breakdown <- create_subtype_table(
    data.frame(Subtype = other_kmt2a_subtypes, 
               Cohort = included_df_modified$Cohort[included_df_modified$Subtype_Merged == "other.KMT2A"]),
    cohort_order
  )
  other_kmt2a_breakdown$Subtype <- paste0("<span style='padding-left:20px;'>└─ ", other_kmt2a_breakdown$Subtype, "</span>")
}

# MECOM breakdown (GATA2;MECOM, MECOM other -> merged as MECOM rearrangement)
mecom_subtypes <- included_df_modified$Subtype_Original[included_df_modified$Subtype_Merged == "MECOM"]
if (length(mecom_subtypes) > 0) {
  mecom_breakdown <- create_subtype_table(
    data.frame(Subtype = mecom_subtypes,
               Cohort = included_df_modified$Cohort[included_df_modified$Subtype_Merged == "MECOM"]),
    cohort_order
  )
  mecom_breakdown$Subtype <- paste0("<span style='padding-left:20px;'>└─ ", mecom_breakdown$Subtype, "</span>")
}

# other.rare.tx breakdown (rare recurring translocations)
rare_tx_subtypes_included <- included_df_modified$Subtype_Original[
  included_df_modified$Subtype_Merged == "other.rare.tx"
]
if (length(rare_tx_subtypes_included) > 0) {
  other_rare_tx_breakdown <- create_subtype_table(
    data.frame(
      Subtype = rare_tx_subtypes_included,
      Cohort  = included_df_modified$Cohort[included_df_modified$Subtype_Merged == "other.rare.tx"]
    ),
    cohort_order
  )
  # Reorder rare translocations to follow the desired clinical order.
  # Only levels present in the data will appear.
  desired_rare_order <- c(
    "PRDM16::RPN1",
    "NPM1::MLF1",
    "KAT6A::CREBBP",
    "RBM15::MRTF1",
    "NUP98::NSD1",
    "NUP98::KDM5A",
    "NUP98, other partners",
    "ETV6::MNX1",
    "PICALM::MLLT10",
    "FUS::ERG",
    "RUNX1::CBFA2T3",
    "CBFA2T3::GLIS2"
  )
  other_rare_tx_breakdown$Subtype <- factor(
    other_rare_tx_breakdown$Subtype,
    levels = c(
      desired_rare_order,
      setdiff(other_rare_tx_breakdown$Subtype, desired_rare_order)
    )
  )
  other_rare_tx_breakdown <- other_rare_tx_breakdown[
    order(other_rare_tx_breakdown$Subtype),
  ]
  other_rare_tx_breakdown$Subtype <- as.character(other_rare_tx_breakdown$Subtype)

  other_rare_tx_breakdown$Subtype <- paste0(
    "<span style='padding-left:20px;'>└─ ",
    other_rare_tx_breakdown$Subtype,
    "</span>"
  )
}

# Build the included table with hierarchical structure
included_table_final <- data.frame()
stripe_groups <- c()  # Track which stripe group each row belongs to
current_stripe <- 1

for (i in 1:nrow(included_table_merged)) {
  subtype <- included_table_merged$Subtype[i]
  # Add the merged class row
  included_table_final <- rbind(included_table_final, included_table_merged[i, ])
  stripe_groups <- c(stripe_groups, current_stripe)
  
  # Add breakdown rows if this is a merged class
  if (subtype == "MDS-related" && exists("mds_breakdown")) {
    included_table_final <- rbind(included_table_final, mds_breakdown)
    # Add same stripe group for all breakdown rows
    stripe_groups <- c(stripe_groups, rep(current_stripe, nrow(mds_breakdown)))
  } else if (subtype == "Other KMT2A rearrangements" && exists("other_kmt2a_breakdown")) {
    included_table_final <- rbind(included_table_final, other_kmt2a_breakdown)
    # Add same stripe group for all breakdown rows
    stripe_groups <- c(stripe_groups, rep(current_stripe, nrow(other_kmt2a_breakdown)))
  } else if (subtype == "MECOM rearrangement" && exists("mecom_breakdown")) {
    included_table_final <- rbind(included_table_final, mecom_breakdown)
    stripe_groups <- c(stripe_groups, rep(current_stripe, nrow(mecom_breakdown)))
  } else if (subtype == "AML with other rare recurring translocations" &&
             exists("other_rare_tx_breakdown")) {
    # Add expanded rows for rare translocations under the merged row
    included_table_final <- rbind(included_table_final, other_rare_tx_breakdown)
    stripe_groups <- c(stripe_groups, rep(current_stripe, nrow(other_rare_tx_breakdown)))
  }
  
  # Increment stripe for next group
  current_stripe <- current_stripe + 1
}

# Create excluded table, with rare recurrent translocations grouped
excluded_df_modified <- excluded_df
excluded_df_modified$Subtype_Merged <- excluded_df$Subtype
excluded_df_modified$Subtype_Merged[
  excluded_df_modified$Subtype %in% rare_tx_subtypes
] <- "other.rare.tx"

excluded_table <- create_subtype_table(
  data.frame(
    Subtype = excluded_df_modified$Subtype_Merged,
    Cohort = excluded_df_modified$Cohort
  ),
  cohort_order
)

# Rename merged rare group for excluded section as well
excluded_table$Subtype <- gsub(
  "other\\.rare\\.tx",
  "AML with other rare recurring translocations",
  excluded_table$Subtype
)

# Add Group column
included_table_final$Group <- "Included"
excluded_table$Group <- "Excluded"

# Add totals per study split by included/excluded
included_totals <- data.frame(
  Subtype = "Total",
  t(colSums(included_table_merged[, cohort_order])),  # Use merged table for correct totals
  Total = sum(included_table_merged$Total),
  Group = "Included",
  stringsAsFactors = FALSE
)
colnames(included_totals) <- c("Subtype", cohort_order, "Total", "Group")

excluded_totals <- data.frame(
  Subtype = "Total",
  t(colSums(excluded_table[, cohort_order])),
  Total = sum(excluded_table$Total),
  Group = "Excluded",
  stringsAsFactors = FALSE
)
colnames(excluded_totals) <- c("Subtype", cohort_order, "Total", "Group")

# Combine all tables
combined_table_with_totals <- rbind(
  included_table_final[, c("Group", "Subtype", cohort_order, "Total")],
  included_totals[, c("Group", "Subtype", cohort_order, "Total")],
  excluded_table[, c("Group", "Subtype", cohort_order, "Total")],
  excluded_totals[, c("Group", "Subtype", cohort_order, "Total")]
)
combined_table_with_totals$Group <- NULL

# Create formatted table with manual striping
table_output <- kable(combined_table_with_totals, 
      format = "html",
      caption = "Table 1: Distribution of Leukemia Subtypes by Cohort",
      row.names = FALSE, escape = FALSE) %>%
  kable_styling(
    bootstrap_options = c("condensed"),
    full_width = FALSE,
    position = "left"
  )

# Apply manual striping for included rows
# Stripe colors for included section - merged classes share stripe with their breakdowns
for (i in 1:length(stripe_groups)) {
  stripe_color <- if (stripe_groups[i] %% 2 == 1) "#f9f9f9" else "white"
  table_output <- table_output %>%
    row_spec(i, background = stripe_color)
}

# Apply striping for excluded rows (standard alternating)
excluded_start <- nrow(included_table_final) + 2
excluded_end <- nrow(combined_table_with_totals) - 1
for (i in excluded_start:excluded_end) {
  row_index <- i - excluded_start + 1
  stripe_color <- if (row_index %% 2 == 1) "#f9f9f9" else "white"
  table_output <- table_output %>%
    row_spec(i, background = stripe_color)
}

# Add section headers and total row styling
table_output <- table_output %>%
  pack_rows("Included", 1, nrow(included_table_final)+1, 
            label_row_css = "background-color: #e6f2ff; font-weight: bold;", indent = F) %>%
  pack_rows("Excluded", excluded_start, nrow(combined_table_with_totals),
            label_row_css = "background-color: #ffe6e6; font-weight: bold;", indent = F) %>%
  add_header_above(c(" " = 1, "Adult Cohorts" = 4, "Pediatric Cohorts" = 3, " " = 1))

table_output
save_kable(table_output, "../writing/tables_new/supplementary_table1.png", zoom = 2)
```

## -- Table 1: Included Subtypes Only

```{r results='asis'}
# Table with only included classes, with MDS-related, other KMT2A, and MECOM
# subtypes expanded into their component subtypes (matching Supplementary Table 1)
# Add a total row (based on merged-class counts to avoid double-counting)
included_totals_simple <- data.frame(
  Subtype = "Total",
  t(colSums(included_table_merged[, cohort_order])),
  Total = sum(included_table_merged$Total),
  stringsAsFactors = FALSE
)
colnames(included_totals_simple) <- c("Subtype", cohort_order, "Total")
# Combine with the hierarchical included table (drop Group column)
# that already contains the split MDS and other KMT2A subtypes
included_simple <- rbind(
  included_table_final[, c("Subtype", cohort_order, "Total")],
  included_totals_simple
)

length(included_table_merged$Subtype)

# Create formatted table
simple_table <- kable(included_simple, 
      format = "html",
      caption = "Table 1: Included Leukemia Subtypes by Cohort",
      row.names = FALSE, 
      escape = FALSE) %>%
  kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(nrow(included_simple), bold = TRUE, background = "#e6f2ff") %>%
  add_header_above(c(" " = 1, "Adult Cohorts" = 4, "Pediatric Cohorts" = 3, " " = 1))

simple_table
save_kable(simple_table, "../writing/tables_new/table1.png", zoom = 2)
```

```{r}
included_simple_perc <- data.frame(class = included_simple$Subtype, perc = included_simple$Total/max(included_simple$Total) * 100, n = included_simple$Total)
included_simple_perc$class <- gsub("<span style='padding-left:20px;'>└─ ", "", included_simple_perc$class)
included_simple_perc$class <- gsub("</span>", "", included_simple_perc$class)

included_simple_perc
```

## -- patients and classes per age catagory
```{r}
# Convert age to years and create 1-year bins
meta$age_years <- meta$age_days / 365.25
max_age <- ceiling(max(meta$age_years, na.rm = TRUE))
age_bins <- seq(0, max_age, by = 1)

# Assign each patient to an age bin
meta$age_bin <- cut(meta$age_years, breaks = age_bins, right = FALSE,
                    labels = paste0(head(age_bins, -1), "-", tail(age_bins, -1)))

# Calculate patients and unique classes per age bin
age_summary <- meta %>%
  filter(!is.na(age_bin)) %>%
  group_by(age_bin) %>%
  summarise(
    n_patients = n(),
    n_classes = n_distinct(classes),
    .groups = "drop"
  ) %>%
  mutate(age_midpoint = as.numeric(gsub("-.*", "", as.character(age_bin))) + 0.5)

# Simple histogram: patients per 1-year age bin
p_patients <- ggplot(meta %>% filter(!is.na(age_years)), aes(x = age_years)) +
  geom_histogram(breaks = age_bins, fill = "#7eb0d5", color = "white", alpha = 0.8) +
  geom_vline(xintercept = 18, linetype = "dashed", color = "gray40") +
  annotate("text", x = 18, y = Inf, label = "18y", vjust = 2, hjust = -0.2, 
           size = 3, color = "gray40") +
  theme_bw() +
  labs(
    title = "Patient Distribution by Age",
    subtitle = "1-year bins",
    x = "Age (years)",
    y = "Number of Patients"
  ) +
  scale_x_continuous(breaks = seq(0, max(age_bins), by = 10)) +
  theme(text = element_text(size = 11))

print(p_patients)
```

```{r}
# Combined plot: patients and unique classes per age bin
# Reshape data for faceted plot
age_summary_long <- age_summary %>%
  tidyr::pivot_longer(
    cols = c(n_patients, n_classes),
    names_to = "metric",
    values_to = "count"
  ) %>%
  mutate(metric = factor(metric, 
                         levels = c("n_patients", "n_classes"),
                         labels = c("Patients", "Unique Classes")))

p_combined <- ggplot(age_summary_long, aes(x = age_midpoint, y = count, fill = metric)) +
  geom_col(color = "white", width = 0.9) +
  geom_vline(xintercept = 18, linetype = "dashed", color = "gray40") +
  facet_wrap(~ metric, scales = "free_y", ncol = 2) +
  scale_fill_manual(values = c("Patients" = "#7eb0d5", "Unique Classes" = "#fd7f6f")) +
  theme_bw() +
  labs(
    title = "Patient and Class Distribution by Age",
    subtitle = sprintf("Total: %d patients, %d unique classes (1-year bins)", 
                       nrow(meta), n_distinct(meta$classes)),
    x = "Age (years)",
    y = "Count"
  ) +
  scale_x_continuous(breaks = seq(0, max(age_bins), by = 10)) +
  theme(
    text = element_text(size = 11),
    legend.position = "none",
    strip.background = element_rect(fill = "gray95"),
    strip.text = element_text(face = "bold")
  )

print(p_combined)

age_summary_long %>% filter(metric == "Unique Classes") %>% summarize(mean(as.numeric(count), na.rm = TRUE))
```

```{r fig.width=8, fig.height=5}
# Alternative: overlaid bar chart with dual y-axis
# Scale factor to align the two metrics visually
scale_factor <- max(age_summary$n_patients) / max(age_summary$n_classes)

p_dual <- ggplot(age_summary, aes(x = age_midpoint)) +
  geom_col(aes(y = n_patients), fill = "#7eb0d5", alpha = 0.7, width = 0.9) +
  geom_line(aes(y = n_classes * scale_factor), color = "#fd7f6f", linewidth = 1.2) +
  geom_point(aes(y = n_classes * scale_factor), color = "#fd7f6f", size = 2.5) +
  geom_vline(xintercept = 18, linetype = "dashed", color = "gray40") +
  scale_y_continuous(
    name = "Number of Patients",
    sec.axis = sec_axis(~ . / scale_factor, name = "Unique Classes")
  ) +
  scale_x_continuous(breaks = seq(0, max(age_bins), by = 10)) +
  theme_bw() +
  labs(
    title = "Patient Count and Class Diversity by Age",
    subtitle = "Bars: patients | Line: unique classes (1-year bins)",
    x = "Age (years)"
  ) +
  theme(
    text = element_text(size = 11),
    axis.title.y.left = element_text(color = "#7eb0d5"),
    axis.title.y.right = element_text(color = "#fd7f6f"),
    axis.text.y.right = element_text(color = "#fd7f6f")
  )

print(p_dual)
```



##  --Performance Boxplots

```{r}
types <- c("cv", "loso")
models <- names(outer_merged$detailed_performance$cv)

# Helper: extract per-fold performance for a given result object and label set
get_results_per_fold <- function(res_obj, label_set) {
  out <- data.frame()
  for (type_i in types) {
    for (model_i in models) {
      folds <- names(res_obj$detailed_performance[[type_i]][[models[1]]])
      for (fold_i in folds) {
        out <- rbind(
          out,
          data.frame(
            type      = type_i,
            model     = model_i,
            fold      = fold_i,
            kappa     = res_obj$detailed_performance[[type_i]][[model_i]][[fold_i]][["kappa"]],
            accuracy  = res_obj$detailed_performance[[type_i]][[model_i]][[fold_i]][["accuracy"]],
            label_set = label_set,
            stringsAsFactors = FALSE
          )
        )
      }
    }
  }
  out
}

# Combine unmerged (full subtype resolution) and merged (collapsed) results
results_per_fold <- rbind(
  get_results_per_fold(outer_unmerged, "Full subtypes"),
  get_results_per_fold(outer_merged,  "Collapsed classes")
)

library(ggplot2)
library(dplyr)
library(tidyr)
library(ComplexHeatmap)
library(circlize)
library(ggplotify)
library(grid)

# Clean model names
results_per_fold$model <- gsub("_", " ", results_per_fold$model)
results_per_fold$model <- gsub("xgboost", "XGBoost", results_per_fold$model)
results_per_fold$model <- gsub("svm", "SVM", results_per_fold$model)
results_per_fold$model <- gsub("neural net", "DNN", results_per_fold$model)
results_per_fold$model <- gsub("Optimized", "Ensemble", results_per_fold$model)

results_per_fold$label_set <- factor(
  results_per_fold$label_set,
  levels = c("Full subtypes", "Collapsed classes")
)

# Colour palette per model (used for outlines / points)
colours <- c("#fd7f6f", "#7eb0d5", "#b2e061", "#bd7ebe", "#ffb55a")
names(colours) <- c("XGBoost", "DNN", "SVM", "OvR Ensemble", "Global Ensemble")

# Subtle fill colours to distinguish full vs collapsed labels
label_fills <- c("Full subtypes" = "#f0f0f0", "Collapsed classes" = "#bdbdbd")

# Compute per-model, per-label-set means for annotations
results_per_fold <- results_per_fold %>%
  group_by(type, model, label_set) %>%
  mutate(mean_kappa = mean(kappa)) %>%
  ungroup()

means_cv <- results_per_fold %>%
  filter(type == "cv") %>%
  group_by(model, label_set) %>%
  summarise(mean_kappa = mean(kappa), .groups = "drop")

means_loso <- results_per_fold %>%
  filter(type == "loso") %>%
  group_by(model, label_set) %>%
  summarise(mean_kappa = mean(kappa), .groups = "drop")

# Order models on the x-axis from low to high median kappa, separately for CV and LOSO
cv_order <- results_per_fold %>%
  filter(type == "cv", label_set == "Collapsed classes") %>%
  group_by(model) %>%
  summarise(median_kappa = mean(kappa), .groups = "drop") %>%
  arrange(median_kappa) %>%
  pull(model)

loso_order <- results_per_fold %>%
  filter(type == "loso", label_set == "Collapsed classes") %>%
  group_by(model) %>%
  summarise(median_kappa = mean(kappa), .groups = "drop") %>%
  arrange(median_kappa) %>%
  pull(model)

results_per_fold$model_cv   <- factor(results_per_fold$model,   levels = cv_order)
results_per_fold$model_loso <- factor(results_per_fold$model, levels = loso_order)

means_cv$model   <- factor(means_cv$model,   levels = cv_order)
means_loso$model <- factor(means_loso$model, levels = loso_order)

box_width <- 0.6

# CV boxplots: per model, side-by-side boxes for full vs collapsed classes
p1 <- ggplot(
  results_per_fold %>% filter(type == "cv"),
  aes(
    x     = model_cv,
    y     = kappa,
    fill  = label_set,
    color = model,
    group = interaction(model, label_set),
    show_legend = FALSE
  )
) +
  geom_boxplot(
    width         = box_width,
    alpha         = 0.9,
    outlier.shape = NA,
    position      = position_dodge(width = 0.7),
    show_legend = FALSE
  ) +
  geom_jitter(
    aes(color = model),
    size     = 1.8,
    alpha    = 0.6,
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.7),
    show_legend = FALSE
  ) +
  geom_text(
    data = means_cv,
    aes(
      x     = model,
      y     = 0.76,
      angle = 45,
      label = sprintf("%.2f", mean_kappa),
      group = interaction(model, label_set),
      color = model
    ),
    position    = position_dodge(width = 0.7),
    inherit.aes = FALSE,
    size        = 3.2
  ) +
  theme_bw() +
  labs(
    title = "Regular nested CV",
    x     = "",
    y     = "Cohen's Kappa",
    fill  = "Label set",
    color = "Model"
  ) +
  theme(
    axis.text.x  = element_text(angle = 45, hjust = 1),
    legend.location = "none"
  ) +
  ylim(0.75, 1) +
  scale_fill_manual(values = label_fills) +
  scale_color_manual(values = colours) + guides(color=FALSE, fill = FALSE)

# LOSO boxplots: same structure as CV
p2 <- ggplot(
  results_per_fold %>% filter(type == "loso"),
  aes(
    x     = model_loso,
    y     = kappa,
    fill  = label_set,
    color = model,
    group = interaction(model, label_set)
  )
) +
  geom_boxplot(
    width         = box_width,
    alpha         = 0.9,
    outlier.shape = NA,
    position      = position_dodge(width = 0.7)
  ) +
  geom_jitter(
    aes(color = model),
    size     = 1.8,
    alpha    = 0.6,
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.7)
  ) +
  geom_text(
    data = means_loso,
    aes(
      x     = model,
      y     = 0.76,
      angle = 45,
      label = sprintf("%.2f", mean_kappa),
      group = interaction(model, label_set),
      color = model
    ),
    position    = position_dodge(width = 0.7),
    inherit.aes = FALSE,
    size        = 3.2
  ) +
  theme_bw() +
  labs(
    title = "LOSO nested CV",
    x     = "",
    y     = "Cohen's Kappa",
    fill  = "Label set",
    color = "Model"
  ) +
  theme(
    axis.text.x  = element_text(angle = 45, hjust = 1),
    legend.box   = "vertical",
    legend.title = element_text(size = 9),
    legend.text  = element_text(size = 8)
  ) +
  ylim(0.75, 1) +
  scale_fill_manual(values = label_fills,guide = guide_legend(reverse = TRUE)) +
  scale_color_manual(values = colours, guide = guide_legend(reverse = TRUE))

library(patchwork)
# Arrange CV and LOSO boxplots vertically and collect a single shared legend
boxplot_figure <- (p1 + p2) + patchwork::plot_annotation(title = "Performance by model and label granularity")
boxplot_figure
```

##  --Performance Table (Mean ± SD)

```{r}

# Summary heatmaps of mean and SD per model for CV and LOSO (including accuracy),
# now with separate columns for full vs collapsed label sets.
summary_stats <- results_per_fold %>%
  group_by(type, model, label_set) %>%
  summarise(
    mean_kappa    = mean(kappa),
    sd_kappa      = sd(kappa),
    mean_accuracy = mean(accuracy),
    sd_accuracy   = sd(accuracy),
    .groups       = "drop"
  )

# Create table with mean (±SD) format directly from summary_stats
# Extract and rename columns for each combination of type and label_set
cv_full <- summary_stats %>%
  filter(type == "cv", label_set == "Full subtypes") %>%
  select(model, 
         cv_kappa_mean_full = mean_kappa, 
         cv_kappa_sd_full = sd_kappa,
         cv_acc_mean_full = mean_accuracy, 
         cv_acc_sd_full = sd_accuracy)

cv_collapsed <- summary_stats %>%
  filter(type == "cv", label_set == "Collapsed classes") %>%
  select(model,
         cv_kappa_mean_collapsed = mean_kappa,
         cv_kappa_sd_collapsed = sd_kappa,
         cv_acc_mean_collapsed = mean_accuracy,
         cv_acc_sd_collapsed = sd_accuracy)

loso_full <- summary_stats %>%
  filter(type == "loso", label_set == "Full subtypes") %>%
  select(model,
         loso_kappa_mean_full = mean_kappa,
         loso_kappa_sd_full = sd_kappa,
         loso_acc_mean_full = mean_accuracy,
         loso_acc_sd_full = sd_accuracy)

loso_collapsed <- summary_stats %>%
  filter(type == "loso", label_set == "Collapsed classes") %>%
  select(model,
         loso_kappa_mean_collapsed = mean_kappa,
         loso_kappa_sd_collapsed = sd_kappa,
         loso_acc_mean_collapsed = mean_accuracy,
         loso_acc_sd_collapsed = sd_accuracy)

# Join all data together
performance_table <- cv_full %>%
  left_join(cv_collapsed, by = "model") %>%
  left_join(loso_full, by = "model") %>%
  left_join(loso_collapsed, by = "model") %>%
  mutate(
    Model = model,
    `CV Kappa (Full)` = sprintf("%.2f ± %.2f", cv_kappa_mean_full, cv_kappa_sd_full),
    `CV Kappa (Collapsed)` = sprintf("%.2f ± %.2f", cv_kappa_mean_collapsed, cv_kappa_sd_collapsed),
    `CV Accuracy (Full)` = sprintf("%.1f%% ± %.1f%%", cv_acc_mean_full * 100, cv_acc_sd_full * 100),
    `CV Accuracy (Collapsed)` = sprintf("%.1f%% ± %.1f%%", cv_acc_mean_collapsed * 100, cv_acc_sd_collapsed * 100),
    `LOSO Kappa (Full)` = sprintf("%.2f ± %.2f", loso_kappa_mean_full, loso_kappa_sd_full),
    `LOSO Kappa (Collapsed)` = sprintf("%.2f ± %.2f", loso_kappa_mean_collapsed, loso_kappa_sd_collapsed),
    `LOSO Accuracy (Full)` = sprintf("%.1f%% ± %.1f%%", loso_acc_mean_full * 100, loso_acc_sd_full * 100),
    `LOSO Accuracy (Collapsed)` = sprintf("%.1f%% ± %.1f%%", loso_acc_mean_collapsed * 100, loso_acc_sd_collapsed * 100)
  ) %>%
  select(Model, `CV Kappa (Full)`, `CV Kappa (Collapsed)`, `CV Accuracy (Full)`, 
         `CV Accuracy (Collapsed)`, `LOSO Kappa (Full)`, `LOSO Kappa (Collapsed)`, 
         `LOSO Accuracy (Full)`, `LOSO Accuracy (Collapsed)`)

# Order by cv_order (reverse order to match original)
performance_table <- performance_table %>%
  mutate(Model = factor(Model, levels = cv_order)) %>%
  arrange(desc(Model)) %>%
  mutate(Model = as.character(Model))

# Display table using kable
performance_table_kable <- kable(
  performance_table,
  caption = "Model Performance Metrics (Mean ± SD)",
  align = c("l", rep("c", 8)),
  col.names = NULL
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    position = "center"
  ) %>%
  column_spec(1, bold = TRUE) %>%
  add_header_above(c(" " = 1, "Full" = 1, "Collapsed" = 1, "Full" = 1, "Collapsed" = 1, "Full" = 1, "Collapsed" = 1, "Full" = 1, "Collapsed" = 1)) %>%
  add_header_above(c(" " = 1, "Kappa (Mean ± SD)" = 2, "Accuracy (Mean ± SD)" = 2, "Kappa (Mean ± SD)" = 2, "Accuracy (Mean ± SD)" = 2)) %>%
  add_header_above(c(" " = 1, "Standard CV" = 4, "LOSO" = 4)) 
  

performance_table_kable
```

# Per-Class Performance Analysis

##  --Per-Class Metrics by Subtype

```{r}
mapping <- c(
    "AML.with.inv.16..t.16.16..CBFB..MYH11" = "CBFB::MYH11",
    "APL..t.15.17..PML..RARA" = "PML::RARA",
    "AML.with.t.6.9..DEK..NUP214" = "DEK::NUP214",
    "AML.with.mutated.NPM1" = "Mutated NPM1",
    "AML.with.t.8.21..RUNX1..RUNX1T1" = "RUNX1::RUNX1T1",
    "AML.with.in.frame.bZIP.CEBPA" = "CEBPA bZIP in-frame",
    "ETV6..MNX1" = "ETV6::MNX1",
    "KAT6A..CREBBP" = "KAT6A::CREBBP",
    "RBM15..MRTF1" = "RBM15::MRTF1",
    "MDS.r" = "MDS-related",
    "AML.with.t.9.11..MLLT3..KMT2A" = "MLLT3::KMT2A",
    "FUS..ERG" = "FUS::ERG",
    "NUP98..NSD1" = "NUP98::NSD1",
    "other.KMT2A" = "Other KMT2A rearrangements",
    "CBFA2T3..GLIS2" = "CBFA2T3::GLIS2",
    "NUP98..KDM5A" = "NUP98::KDM5A",
    "NUP98..other.partners" = "NUP98, other partners",
    # MDS-related subtypes (all map to MDS-related)
    "AML.with.MDS.related.gene.mutations" = "MDS-related, gene mutations",
    "AML.with.MDS.related.cytogenetic.abnormalities" = "MDS-related, cytogenetic abnormalities",
    "AML.with.mutated.TP53" = "MDS-related, TP53 mutated",
    # MECOM: merged and unmerged (GATA2;MECOM, MECOM other)
    "MECOM" = "MECOM rearrangement",
    "MECOM.fusion" = "MECOM rearrangement",
    "MECOM.rearrangement" = "MECOM rearrangement",
    "GATA2.MECOM" = "GATA2;MECOM",
    "MECOM.other" = "MECOM, other",
    "MECOM..other" = "MECOM, other",
    # Additional subtypes that may appear in results
    "AML.with.t.9.22..BCR..ABL1" = "BCR::ABL1",
    "BCR..ABL1" = "BCR::ABL1",
    "PRDM16..RPN1" = "PRDM16::RPN1",
    "NPM1..MLF1" = "NPM1::MLF1",
    "PICALM..MLLT10" = "PICALM::MLLT10",
    "RUNX1..CBFA2T3" = "RUNX1::CBFA2T3",
    # KMT2A variants
    "KMT2A..ELL" = "KMT2A::ELL",
    "KMT2A..MLLT1" = "KMT2A::MLLT1",
    "KMT2A..MLLT10" = "KMT2A::MLLT10",
    "KMT2A..MLLT4" = "KMT2A::MLLT4",
    "KMT2A..MLLT4....AFDN." = "KMT2A::MLLT4",
    "KMT2A..SEPT6" = "KMT2A::SEPT6"
  )

# Order matching the table with all included subtypes (desired_included_order)
# This matches the ordering used in Supplementary Table 1
order_AML <- c(
    "PML::RARA",                                    # APL, t(15;17)/PML::RARA
    "RUNX1::RUNX1T1",                              # AML with t(8;21)/RUNX1::RUNX1T1
    "CBFB::MYH11",                                 # AML with inv(16)/t(16;16)/CBFB::MYH11
    "MLLT3::KMT2A",                                # AML with t(9;11)/MLLT3::KMT2A
    "Other KMT2A rearrangements",                  # Other KMT2A rearrangements
    # Other KMT2A subtypes
    "KMT2A::ELL",
    "KMT2A::MLLT1",
    "KMT2A::MLLT10",
    "KMT2A::MLLT4",
    "KMT2A::SEPT6",
    "DEK::NUP214",                                 # AML with t(6;9)/DEK::NUP214
    "MECOM rearrangement",                         # MECOM rearrangement (merged)
    "GATA2;MECOM",                                 # MECOM unmerged
    "MECOM, other",                                # MECOM unmerged
    "BCR::ABL1",                                   # AML with t(9;22)/BCR::ABL1
    "PRDM16::RPN1",
    "NPM1::MLF1",
    "KAT6A::CREBBP",
    "RBM15::MRTF1",
    "NUP98::NSD1",                                 # NUP98::NSD1
    "NUP98::KDM5A",
    "NUP98, other partners",
    "ETV6::MNX1",
    "PICALM::MLLT10",
    "FUS::ERG",
    "RUNX1::CBFA2T3",
    "CBFA2T3::GLIS2",
    "Mutated NPM1",                                # AML with mutated NPM1
    "CEBPA bZIP in-frame",                         # AML with in-frame bZIP CEBPA
    "MDS-related",          # MDS-related
    "MDS-related, gene mutations",
    "MDS-related, TP53 mutated",
    "MDS-related, cytogenetic abnormalities"

  )
# Global_Optimized is best
fix_names <- function(x){
  # Map verbose labels to concise AML names; preserves factor inputs by relabeling levels

  if (is.factor(x)) {
    new_levels <- vapply(levels(x), function(l) {
      val <- unname(mapping[l])
      if (is.na(val)) l else val
    }, character(1))
    levels(x) <- new_levels
    return(x)
  } else {
    y <- vapply(as.character(x), function(l) {
      val <- unname(mapping[l])
      if (is.na(val)) l else val
    }, character(1))
    return(y)
  }
}

make_cm_fig <- function(res_obj, type, label_set) {
  # All per-class summaries for a given result object (full vs collapsed)
  per_class <- res_obj$per_class_summaries[[type]]
  per_class <- per_class[per_class$Model == "OvR_Ensemble", ]
  rownames(per_class) <- gsub("Class: ", "", per_class$Class)

  # Apply pretty names to per-class row labels
  rownames(per_class) <- fix_names(rownames(per_class))

  # Get subtypes that exist in both per_class and order_AML, preserving order_AML order
  available_subtypes <- intersect(order_AML, rownames(per_class))
  # if (length(available_subtypes) < lenght(order_AML)){
  #   print("SOmething wrong with n classes")
  # }
  per_class <- per_class[available_subtypes, c(
    "Mean_F1",
    "Mean_Sensitivity",
    "Mean_Specificity",
    "Mean_Precision",
    "Mean_Recall"
  )]
  colnames(per_class) <- gsub("Mean_", "", colnames(per_class))
  print(sum(per_class$F1 > 0.8, na.rm = T))
  print(sum(per_class$F1 > 0.9, na.rm = T))
  #to_remove <- apply(per_class, 1, function(x) any(is.na(x)))
  #per_class <- per_class[!to_remove, ]
per_class <- apply(per_class, 2, function(x) round(x, 2))
  normalise <- function(x) {
    (x - min(x)) / (max(x) - min(x))
  }
  per_class_norm <- apply(per_class, 2, normalise)

  # Define palette for per-class metrics (white to green for performance)
  palette_fun <- circlize::colorRamp2(c(0, 1), c("white", "green4"))

  ht_per_class <- Heatmap(
    per_class,
    name              = NULL,
    row_names_side    = "left",
    column_names_side = "top",
    row_names_gp      = gpar(fontsize = 10),
    column_names_gp   = gpar(fontsize = 10),
    cluster_rows      = FALSE,
    cluster_columns   = FALSE,
    rect_gp           = gpar(col = "#d9d9d9", lwd = 0.5),
    cell_fun = function(j, i, x, y, width, height, fill) {
      val      <- per_class[i, j]
      norm_val <- per_class_norm[i, j]
      grid.rect(
        x,
        y,
        width  = width,
        height = height,
        gp     = gpar(
          fill = palette_fun(val),
          col  = "#d9d9d9",
          lwd  = 0.5
        )
      )
      grid.text(sprintf("%.2f", val), x, y, gp = gpar(fontsize = 11))
    },
    show_heatmap_legend  = FALSE,
    column_names_rot     = 45,
    column_names_centered = FALSE,
    column_title         = "Performance measure",
    column_title_gp      = gpar(fontface = "bold"),
    row_title            = "Reference class",
    row_title_gp         = gpar(fontface = "bold")
  )

  # Confusion matrices across folds for the same result object
  conf_table <- list()
  for (fold in names(res_obj$detailed_performance[[type]]$Global_Optimized)) {
      conf_table[[fold]] <- res_obj$detailed_performance[[type]]$OvR_Ensemble[[fold]]$confusion_matrix$table
  }

  # Get all unique classes across all folds
  all_classes <- unique(unlist(lapply(conf_table, function(x) c(rownames(x), colnames(x)))))

  # Pad each confusion matrix to have the same dimensions
  conf_table_padded <- lapply(conf_table, function(mat) {
    padded_mat <- matrix(
      0,
      nrow    = length(all_classes),
      ncol    = length(all_classes),
      dimnames = list(all_classes, all_classes)
    )
    padded_mat[rownames(mat), colnames(mat)] <- mat
    padded_mat
  })

  total_conf <- Reduce("+", conf_table_padded)
  total_conf <- t(total_conf)

  # Apply pretty names to confusion matrix axes and reorder to match per_class
  rownames(total_conf) <- fix_names(rownames(total_conf))
  colnames(total_conf) <- fix_names(colnames(total_conf))
  total_conf <- total_conf[rownames(per_class), rownames(per_class)]
  
  total_conf_norm <- apply(total_conf, 2, normalise)

  total_conf_palette_fun <- circlize::colorRamp2(c(0, 1), c("white", "purple3"))
  ht_table <- Heatmap(
    total_conf,
    name              = NULL,
    row_names_side    = "left",
    column_names_side = "top",
    row_names_gp      = gpar(fontsize = 10),
    column_names_gp   = gpar(fontsize = 10),
    cluster_rows      = FALSE,
    cluster_columns   = FALSE,
    rect_gp           = gpar(col = "#d9d9d9", lwd = 0.5),
    cell_fun = function(j, i, x, y, width, height, fill) {
      val      <- total_conf[i, j]
      norm_val <- total_conf_norm[i, j]
      grid.rect(
        x,
        y,
        width  = width,
        height = height,
        gp     = gpar(
          fill = total_conf_palette_fun(norm_val),
          col  = "#d9d9d9",
          lwd  = 0.5
        )
      )
      grid.text(sprintf("%.0f", val), x, y, gp = gpar(fontsize = 11))
    },
    show_heatmap_legend  = FALSE,
    column_names_rot     = 45,
    column_names_centered = FALSE,
    column_title         = "Predicted class",
    column_title_gp      = gpar(fontface = "bold"),
    row_title            = "Reference class",
    row_title_gp         = gpar(fontface = "bold")
  )

  ht_per_class + ht_table
}

# Build class-performance + confusion-matrix composites for both full and collapsed label sets
cm_fig_cv_full       <- make_cm_fig(outer_unmerged, "cv",   "Full subtypes")
cm_fig_loso_full     <- make_cm_fig(outer_unmerged, "loso", "Full subtypes")
cm_fig_cv_collapsed  <- make_cm_fig(outer_merged,   "cv",   "Collapsed classes")
cm_fig_loso_collapsed <- make_cm_fig(outer_merged,  "loso", "Collapsed classes")

# Wrap as ggplot objects with informative titles
cm_fig_cv_full_plot <- ggplotify::as.ggplot(function() {
  ComplexHeatmap::draw(
    cm_fig_cv_full,
    padding        = unit(c(5, 3, 5, 50), "mm"),
    column_title   = "CV: Class performance (Full subtypes)",
    column_title_gp = gpar(fontsize = 12, fontface = "bold")
  )
})

cm_fig_cv_collapsed_plot <- ggplotify::as.ggplot(function() {
  ComplexHeatmap::draw(
    cm_fig_cv_collapsed,
    padding        = unit(c(5, 3, 5, 50), "mm"),
    column_title   = "CV: Class performance (Collapsed classes)",
    column_title_gp = gpar(fontsize = 12, fontface = "bold")
  )
})

cm_fig_loso_full_plot <- ggplotify::as.ggplot(function() {
  ComplexHeatmap::draw(
    cm_fig_loso_full,
    padding        = unit(c(5, 3, 5, 50), "mm"),
    column_title   = "LOSO: Class performance (Full subtypes)",
    column_title_gp = gpar(fontsize = 12, fontface = "bold")
  )
})

cm_fig_loso_collapsed_plot <- ggplotify::as.ggplot(function() {
  ComplexHeatmap::draw(
    cm_fig_loso_collapsed,
    padding        = unit(c(5, 3, 5, 50), "mm"),
    column_title   = "LOSO: Class performance (Collapsed classes)",
    column_title_gp = gpar(fontsize = 12, fontface = "bold")
  )
})
```
```{r}
library(egg)

# Arrange the four class-performance panels (CV/LOSO × full/collapsed)
figure2_egg <- ggarrange(
  cm_fig_cv_full_plot,      cm_fig_cv_collapsed_plot,
  cm_fig_loso_full_plot,    cm_fig_loso_collapsed_plot,
  ncol  = 2, nrow = 2,
  labels = c("A", "B", "C", "D"),
  label.args = list(gp = gpar(fontface = "bold", cex = 1))
)

ggsave(
  "../writing/figures_new/figures_class_perf.png",
  figure2_egg,
  width  = 400*2,
  height = 300*2,
  units  = "mm"
)
```

## -- Three-Panel Figure: F1 Comparison and Error Analysis

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(ComplexHeatmap)
library(circlize)
library(grid)
library(gridExtra)

# Panel A: F1 Scores Comparison (Full vs Merged)
# Extract F1 scores for full and merged classes
f1_full <- outer_unmerged$per_class_summaries$cv %>%
  filter(Model == "OvR_Ensemble") %>%
  mutate(Class = gsub("Class: ", "", Class),
         Class = fix_names(Class),
         Type = "Full",
         F1 = Mean_F1) %>%
  select(Class, F1, Type)

f1_merged <- outer_merged$per_class_summaries$cv %>%
  filter(Model == "OvR_Ensemble") %>%
  mutate(Class = gsub("Class: ", "", Class),
         Class = fix_names(Class),
         Type = "Merged",
         F1 = Mean_F1) %>%
  select(Class, F1, Type)

# Combine and create comparison data
f1_comparison <- bind_rows(f1_full, f1_merged)

# Define merge mapping for connecting lines (full -> merged class)
merge_mapping <- list(
  "Other KMT2A rearrangements" = c("KMT2A::ELL", "KMT2A::MLLT10", "KMT2A::MLLT4", "KMT2A::SEPT6"),
  "MDS-related" = c("MDS-related, gene mutations", "MDS-related, cytogenetic abnormalities", "MDS-related, TP53 mutated", "MDS.r"),
  "MECOM rearrangement" = c("GATA2;MECOM", "MECOM, other")
)

# Create category groups for coloring
f1_comparison <- f1_comparison %>%
  mutate(
    Category = case_when(
      Class %in% c("PML::RARA", "RUNX1::RUNX1T1", "CBFB::MYH11", "DEK::NUP214", "MLLT3::KMT2A") ~ "Major rearrangements",
      Class %in% c("Other KMT2A rearrangements", "KMT2A::ELL", "KMT2A::MLLT10", "KMT2A::MLLT1","KMT2A::MLLT4", "KMT2A::SEPT6") ~ "Other KMT2A rearrangements",
      Class %in% c("MECOM rearrangement", "GATA2;MECOM", "MECOM, other") ~ "MECOM rearrangement",
      Class %in% c("MDS-related", "MDS-related, gene mutations", "MDS-related, cytogenetic abnormalities", "MDS-related, TP53 mutated", "MDS.r") ~ "MDS-related",
      Class %in% c("Mutated NPM1", "CEBPA bZIP in-frame") ~ "Major mutations",
      Class %in% c("PRDM16::RPN1", "NPM1::MLF1", "KAT6A::CREBBP", "RBM15::MRTF1", "ETV6::MNX1", "PICALM::MLLT10", "FUS::ERG", "RUNX1::CBFA2T3", "CBFA2T3::GLIS2", "NUP98::NSD1", "NUP98::KDM5A", "NUP98, other partners") ~ "Rare translocations",
      TRUE ~ "Other"
    )
  )

# For Panel A: show only Full dot for classes unchanged between Full and Merged; merged groups keep Merged point
classes_in_full <- unique(f1_comparison$Class[f1_comparison$Type == "Full"])
f1_comparison_plot <- f1_comparison %>%
  filter(Type == "Full" | (Type == "Merged" & !(Class %in% classes_in_full)))

# Order classes consistently; include any class in the data so none become NA (unnamed level)
all_classes_ordered <- order_AML
extra_classes <- setdiff(unique(as.character(f1_comparison_plot$Class)), all_classes_ordered)
if (length(extra_classes) > 0) all_classes_ordered <- c(all_classes_ordered, extra_classes)

f1_comparison_plot$Class <- factor(f1_comparison_plot$Class, levels = rev(all_classes_ordered))

# Create color palette for categories
category_colors <- c(
  "Major rearrangements" = "#1f77b4",
  "Other KMT2A rearrangements" = "#ff7f0e",
  "MECOM rearrangement" = "#17becf",
  "Major mutations" = "#2ca02c",
  "Rare translocations" = "#9467bd",
  "MDS-related" = "#d62728"
)

# Panel A: F1 Comparison Plot
library(ggtext)

# Get the actual y-axis levels as they appear on the plot (bottom to top)
axis_class_levels <- levels(f1_comparison_plot$Class)
# Filter to only classes present in the plot data
axis_class_levels <- axis_class_levels[axis_class_levels %in% unique(f1_comparison_plot$Class)]

# Create axis text colors based on category (MDS-related, KMT2A, MECOM get their group colors)
axis_text_colors <- sapply(axis_class_levels, function(cls) {
  if (cls %in% c("MDS-related", "MDS-related, gene mutations", 
                 "MDS-related, cytogenetic abnormalities", "MDS-related, TP53 mutated", "MDS.r")) {
    "#d62728"  # MDS-related color
  } else if (cls %in% c("Other KMT2A rearrangements", 
                        "KMT2A::ELL", "KMT2A::MLLT1", "KMT2A::MLLT10", "KMT2A::MLLT4", 
                        "KMT2A::SEPT6", "KMT2A::SEPT9")) {
    "#ff7f0e"  # KMT2A color
  } else if (cls %in% c("MECOM rearrangement", "GATA2;MECOM", "MECOM, other")) {
    "#17becf"  # MECOM color
  } else {
    "black"
  }
})

# Create labels with bold formatting for merged classes
# Replace ";" with "::" in labels so element_markdown does not parse semicolon as CSS (e.g. GATA2;MECOM -> GATA2::MECOM)
axis_labels <- sapply(axis_class_levels, function(cls) {
  safe_cls <- gsub(";", ", ", cls, fixed = TRUE)
  if (cls %in% c("Other KMT2A rearrangements", "MDS-related", "MECOM rearrangement")) {
    paste0("**", safe_cls, "**")
  } else {
    safe_cls
  }
})
names(axis_labels) <- axis_class_levels

panel_a <- ggplot(f1_comparison_plot, aes(x = F1, y = Class, color = Category, shape = Type)) +
  geom_point(size = 2.5, alpha = 0.8, position = position_dodge(width = 0.5)) +
  geom_vline(xintercept = 0.8, linetype = "dashed", color = "gray50", alpha = 0.5) +
  scale_color_manual(values = category_colors, name = "Category") +
  scale_shape_manual(values = c("Merged" = 17, "Full" = 16), name = "Type") +
  scale_y_discrete(labels = axis_labels) +
  labs(
    x = "F1 Score",
    y = "",
    title = "A. F1 Scores: Full vs Merged Classes"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    axis.text.y = element_markdown(size = 8, color = axis_text_colors),
    axis.text.x = element_text(size = 9),
    legend.position = "right",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 9)
  ) +
  xlim(0, 1.0)
panel_a
```
```{r}
# Panel B & C: Error Heatmaps for KMT2A and MDS groups
# Function to extract and prepare confusion matrix for specific groups
get_error_heatmap <- function(res_obj, type, label_set, focus_groups) {
  # Get confusion matrices across folds
  conf_table <- list()
  
  # Determine model name - check what's available
  model_name <- "OvR_Ensemble"
  
  if (is.null(model_name) || !model_name %in% names(res_obj$detailed_performance[[type]])) {
    return(NULL)
  }
  
  for (fold in names(res_obj$detailed_performance[[type]][[model_name]])) {
    cm_obj <- res_obj$detailed_performance[[type]][[model_name]][[fold]]$confusion_matrix
    if (!is.null(cm_obj) && !is.null(cm_obj$table)) {
      conf_table[[fold]] <- cm_obj$table
    }
  }
  
  if (length(conf_table) == 0) {
    return(NULL)
  }
  
  # Get all unique classes
  all_classes <- unique(unlist(lapply(conf_table, function(x) c(rownames(x), colnames(x)))))
  
  if (length(all_classes) == 0) {
    return(NULL)
  }
  
  # Pad each confusion matrix
  conf_table_padded <- lapply(conf_table, function(mat) {
    padded_mat <- matrix(
      0,
      nrow = length(all_classes),
      ncol = length(all_classes),
      dimnames = list(all_classes, all_classes)
    )
    padded_mat[rownames(mat), colnames(mat)] <- mat
    padded_mat
  })
  
  # Sum across folds
  total_conf <- Reduce("+", conf_table_padded)
  total_conf <- t(total_conf)
  
  # Apply pretty names
  rownames(total_conf) <- fix_names(rownames(total_conf))
  colnames(total_conf) <- fix_names(colnames(total_conf))
  
  # Filter to focus groups - show square matrix of focus classes
  focus_classes <- intersect(focus_groups, rownames(total_conf))
  
  if (length(focus_classes) == 0) {
    return(NULL)
  }
  
  # Show square matrix: focus classes as both rows and columns
  # This shows errors between the focus classes themselves
  conf_focused <- total_conf[focus_classes, focus_classes, drop = FALSE]

  # Check for off-diagonal errors
  off_diag <- conf_focused
  diag(off_diag) <- 0
  
  # Calculate row percentages (error rates)
  row_sums <- rowSums(conf_focused)
  conf_percent <- sweep(conf_focused, 1, row_sums, "/") * 100
  conf_percent[is.nan(conf_percent)] <- 0
  
  return(list(
    counts = conf_focused,
    percentages = conf_percent,
    label_set = label_set
  ))
}

# Define focus groups for KMT2A, MDS, and MECOM
# Panel B: Full classes shown separately
kmt2a_mllt3 <- "MLLT3::KMT2A"
kmt2a_other_full <- c("KMT2A::ELL", "KMT2A::MLLT10", "KMT2A::MLLT1", "KMT2A::MLLT4", "KMT2A::SEPT6")
mds_groups_full <- c("MDS-related, gene mutations", "MDS-related, TP53 mutated", "MDS-related, cytogenetic abnormalities")
mecom_groups_full <- c("GATA2;MECOM", "MECOM, other")

# Panel C: KMT2A::MLLT3 separate, other KMT2A combined, MDS combined, MECOM combined
kmt2a_other_merged <- "Other KMT2A rearrangements"
mds_merged <- "MDS-related"
mecom_merged <- "MECOM rearrangement"

# Combine groups for each panel
focus_groups_panel_b <- c(kmt2a_mllt3, kmt2a_other_full, mds_groups_full, mecom_groups_full)
focus_groups_panel_c <- c(kmt2a_mllt3, kmt2a_other_merged, mds_merged, mecom_merged)

# Get error heatmaps from appropriate result objects
# Panel B uses full classes (outer_unmerged), Panel C uses collapsed classes (outer_merged)
error_panel_b <- get_error_heatmap(outer_unmerged, "cv", "Full", focus_groups_panel_b)
error_panel_c <- get_error_heatmap(outer_merged, "cv", "Collapsed", focus_groups_panel_c)

# Create heatmap function with group annotations
create_error_heatmap <- function(error_data, title, other_kmt2a_classes, mds_classes, mecom_classes = NULL) {
  if (is.null(error_data)) {
    return(NULL)
  }
  
  conf_percent <- error_data$percentages
  
  # Create annotation for groups: MLLT3::KMT2A, Other KMT2A, MDS-related, MECOM
  group_anno <- sapply(rownames(conf_percent), function(name) {
    if (name == "MLLT3::KMT2A") {
      return("KMT2A::MLLT3")
    } else if (name %in% other_kmt2a_classes) {
      return("Other KMT2A")
    } else if (name %in% mds_classes) {
      return("MDS-related")
    } else if (!is.null(mecom_classes) && name %in% mecom_classes) {
      return("MECOM rearrangement")
    } else {
      return("Other")
    }
  })
  
  group_levels <- c("KMT2A::MLLT3", "Other KMT2A", "MDS-related")
  group_cols <- c("KMT2A::MLLT3" = "#1f77b4", "Other KMT2A" = "#ff7f0e", "MDS-related" = "#d62728")
  if (!is.null(mecom_classes)) {
    group_levels <- c(group_levels, "MECOM rearrangement")
    group_cols <- c(group_cols, "MECOM rearrangement" = "#17becf")
  }
  group_anno <- factor(group_anno, levels = group_levels)
  
  ha <- rowAnnotation(
    Group = group_anno,
    col = list(Group = group_cols),
    annotation_legend_param = list(title = "Group", title_gp = gpar(fontsize = 9))
  )
  
  # Color function for percentages (white to purple for errors)
  col_fun <- colorRamp2(c(0, 100), c("white", "purple3"))
  
  ht <- Heatmap(
    conf_percent,
    name = "n samples",
    col = col_fun,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    row_names_side = "left",
    column_names_side = "top",
    row_names_gp = gpar(fontsize = 9),
    column_names_gp = gpar(fontsize = 9),
    cell_fun = function(j, i, x, y, width, height, fill) {
      val <- conf_percent[i, j]
      count <- error_data$counts[i, j]
      grid.rect(x, y, width, height, gp = gpar(fill = fill, col = "#d9d9d9", lwd = 0.5))
      if (i == j) {
        # Diagonal: correct predictions
        grid.text(sprintf("%.0f", count), x, y, gp = gpar(fontsize = 8, fontface = "bold"))
      } else if (count > 0) {
        # Off-diagonal: errors
        grid.text(sprintf("%.0f", count), x, y, gp = gpar(fontsize = 7))
      }
    },
    column_title = title,
    column_title_gp = gpar(fontsize = 11, fontface = "bold"),
    left_annotation = ha,
    heatmap_legend_param = list(
      title = "n samples",
      title_gp = gpar(fontsize = 9),
      labels_gp = gpar(fontsize = 8)
    )
  )
  
  return(ht)
}

# Create Panel B: Full classes (MLLT3::KMT2A separate, other KMT2A separate, MDS separate, MECOM separate)
create_error_heatmap(
  error_panel_b, 
  "B. Error Analysis: Full Classes",
  other_kmt2a_classes = kmt2a_other_full,
  mds_classes = mds_groups_full,
  mecom_classes = mecom_groups_full
)

# Create Panel C: Collapsed classes (MLLT3::KMT2A separate, Other KMT2A combined, MDS combined, MECOM combined)
create_error_heatmap(
  error_panel_c, 
  "C. Error Analysis: Collapsed Classes",
  other_kmt2a_classes = kmt2a_other_merged,
  mds_classes = mds_merged,
  mecom_classes = mecom_merged
)
```



## -- Faceted Confusion Matrices per Fold


```{r}
# Function to create faceted confusion matrices per fold
make_faceted_cm <- function(type) {
  # Get all folds for this type
  folds <- names(outer_merged$detailed_performance[[type]]$OvR_Ensemble)
  # Per class
  per_class_list <- list()
  # Collect confusion matrices for all folds
  conf_matrices <- list()
  for (fold in folds) {
    per_class <- outer_unmerged$detailed_performance[[type]]$svm[[fold]]$confusion_matrix$byClass
    rownames(per_class) <- gsub("Class: ", "", rownames(per_class))
    rownames(per_class) <- fix_names(rownames(per_class))
    
    cm <- outer_unmerged$detailed_performance[[type]]$svm[[fold]]$confusion_matrix$table
    cm <- t(cm)  # Transpose to match convention (rows = reference, cols = prediction)
    
    
    # Apply pretty names
    rownames(cm) <- fix_names(rownames(cm))
    colnames(cm) <- fix_names(colnames(cm))
    
    # Reorder to match standard order
    all_classes <- intersect(order_AML, unique(c(
      rownames(per_class))))
    per_class <- per_class[all_classes, c("F1", "Sensitivity", "Specificity", "Precision", "Recall")]
    #to_remove <- apply(per_class, 1, function(x)
    #  any(is.na(x)))
    #per_class <- per_class[!to_remove, ]
    cm <- cm[rownames(per_class), rownames(per_class), drop = FALSE]
    
    normalise <- function(x) {
      return((x - min(x)) / (max(x) - min(x)))
    }
    per_class_norm <- apply(per_class, 2, normalise)
    
    # Normalize for color scale
    cm_norm <- apply(cm, 2, function(x) {
      if (max(x, na.rm = TRUE) > 0) {
        (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
      } else {
        x
      }
    })
    
    conf_matrices[[fold]] <- list(matrix = cm,
                                  normalized = cm_norm,
                                  per_class = per_class,
                                  per_class_norm = per_class_norm,
                                  fold_name = fold)
  }
  
  # Define palette for confusion matrices (white to purple)
  total_conf_palette_fun <- circlize::colorRamp2(c(0, 1), c("white", "purple3"))
  
  # Define palette for per-class metrics (white to green for performance)
  palette_fun <- circlize::colorRamp2(c(0, 1), c("white", "green4"))
  
  # Create individual heatmaps for each fold
  heatmaps <- list()
  
  for (fold in folds) {
    cm_data <- conf_matrices[[fold]]
    per_class <- cm_data$per_class
    per_class_norm <- cm_data$per_class_norm
    ht_per_class <- Heatmap(
      per_class,
      name = NULL,
      row_names_side = "left",
      column_names_side = "top",
      row_names_gp = gpar(fontsize = 10),
      column_names_gp = gpar(fontsize = 10),
      cluster_rows = FALSE,
      cluster_columns = FALSE,
      rect_gp = gpar(col = "#d9d9d9", lwd = 0.5),
      cell_fun = function(j, i, x, y, width, height, fill) {
        val <- per_class[i, j]
        norm_val <- per_class_norm[i, j]
        grid.rect(
          x,
          y,
          width = width,
          height = height,
          gp = gpar(
            fill = palette_fun(val),
            col = "#d9d9d9",
            lwd = 0.5
          )
        )
        grid.text(sprintf("%.2f", val), x, y, gp = gpar(fontsize = 11))
      },
      show_heatmap_legend = FALSE,
      column_names_rot = 45,
      column_names_centered = F,
      column_title = "Performance measure",
      column_title_gp = gpar(fontface = "bold"),
      row_title = "Reference class",
      row_title_gp = gpar(fontface = "bold")
    )
    ht <- Heatmap(
      cm_data$matrix,
      name = NULL,
      row_names_side = "left",
      column_names_side = "top",
      row_names_gp = gpar(fontsize = 8),
      column_names_gp = gpar(fontsize = 8),
      cluster_rows = FALSE,
      cluster_columns = FALSE,
      rect_gp = gpar(col = "#d9d9d9", lwd = 0.3),
      cell_fun = function(j, i, x, y, width, height, fill) {
        val <- cm_data$matrix[i, j]
        norm_val <- cm_data$normalized[i, j]
        grid.rect(
          x,
          y,
          width = width,
          height = height,
          gp = gpar(
            fill = total_conf_palette_fun(norm_val),
            col = "#d9d9d9",
            lwd = 0.3
          )
        )
        if (val > 0) {
          grid.text(sprintf("%.0f", val), x, y, gp = gpar(fontsize = 8))
        }
      },
      show_heatmap_legend = FALSE,
      column_names_rot = 45,
      column_names_centered = FALSE,
      column_title = paste0("Predicted class for fold: ", cm_data$fold_name),
      column_title_gp = gpar(fontsize = 10, fontface = "bold"),
      row_title = "Reference",
      row_title_gp = gpar(fontsize = 9, fontface = "bold")
    )
    
    heatmaps[[fold]] <- ht_per_class+ht
  }
  
  return(heatmaps)
}

# Create faceted confusion matrices for CV
cat("Creating faceted confusion matrices for CV...\n")
cm_heatmaps_cv <- make_faceted_cm("cv")

# Create faceted confusion matrices for LOSO
cat("Creating faceted confusion matrices for LOSO...\n")
cm_heatmaps_loso <- make_faceted_cm("loso")

# Draw CV confusion matrices
cat("Saving CV faceted confusion matrices...\n")
png(
  "../writing/figures_new/confusion_matrix_cv_faceted.png",
  width = 1500,
  height = 700*5,
  pointsize = 10,
  res = 150
)

# Arrange CV heatmaps
ncol_cv <- 1
nrow_cv <- ceiling(length(cm_heatmaps_cv) / ncol_cv)

# Create a layout using grid
grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow_cv, ncol_cv)))

for (i in 1:length(cm_heatmaps_cv)) {
  row <- ((i - 1) %/% ncol_cv) + 1
  col <- ((i - 1) %% ncol_cv) + 1
  
  pushViewport(viewport(layout.pos.row = row, layout.pos.col = col))
  ComplexHeatmap::draw(cm_heatmaps_cv[[i]], newpage = FALSE)
  popViewport()
}

popViewport()
dev.off()

# Draw LOSO confusion matrices
cat("Saving LOSO faceted confusion matrices...\n")
png(
  "../writing/figures_new/confusion_matrix_loso_faceted.png",
  width = 1500,
  height = 700*7,
  pointsize = 10,
  res = 150
)

# Arrange LOSO heatmaps (7 folds)
ncol_loso <- 1
nrow_loso <- ceiling(length(cm_heatmaps_loso) / ncol_loso)

grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow_loso, ncol_loso)))

for (i in 1:length(cm_heatmaps_loso)) {
  row <- ((i - 1) %/% ncol_loso) + 1
  col <- ((i - 1) %% ncol_loso) + 1
  
  pushViewport(viewport(layout.pos.row = row, layout.pos.col = col))
  ComplexHeatmap::draw(cm_heatmaps_loso[[i]], newpage = FALSE)
  popViewport()
}

popViewport()
dev.off()

cat("Faceted confusion matrices saved!\n")
```

# Performance stratification

##  --Performance by Fold

```{r}
# Define pediatric cohorts for color coding
pediatric_cohorts <- c("AAML0531", "AAML1031", "AAML03P1")

# Helper function to order folds by mean kappa and create color vector for x-axis labels
prepare_fold_data <- function(data_subset) {
  fold_means <- data_subset %>% filter(model == "OvR Ensemble") %>%
    group_by(fold) %>%
    summarise(mean_kappa = mean(kappa), .groups = "drop") %>%
    arrange(mean_kappa)
  
  data_subset$fold <- factor(data_subset$fold, levels = fold_means$fold)
  fold_order <- levels(data_subset$fold)
  label_colors <- ifelse(fold_order %in% pediatric_cohorts, "#fd7f6f", "#7eb0d5")
  
  return(list(data = data_subset, label_colors = label_colors))
}

# CV: Full subtypes
cv_full_data <- results_per_fold[results_per_fold$type == "cv" & results_per_fold$label_set == "Full subtypes", ]
cv_full_data$fold <- factor(cv_full_data$fold, levels = sort(unique(as.character(cv_full_data$fold))))
pa <- ggplot(cv_full_data, aes(x = fold, y = kappa, color = model, group = model)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  labs(title = "CV: Full subtypes", x = "Fold", y = "Cohen's Kappa", color = "Model") + 
  scale_color_manual(values = colours, breaks = rev(names(colours)), limits = rev(names(colours))) + 
  ylim(0.75, 1)

# CV: Collapsed classes
cv_collapsed_data <- results_per_fold[results_per_fold$type == "cv" & results_per_fold$label_set == "Collapsed classes", ]
cv_collapsed_data$fold <- factor(cv_collapsed_data$fold, levels = sort(unique(as.character(cv_collapsed_data$fold))))
pb <- ggplot(cv_collapsed_data, aes(x = fold, y = kappa, color = model, group = model)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  labs(title = "CV: Collapsed classes", x = "Fold", y = "Cohen's Kappa", color = "Model") + 
  scale_color_manual(values = colours, breaks = rev(names(colours)), limits = rev(names(colours))) + 
  ylim(0.75, 1)

# LOSO: Full subtypes - order by mean kappa
loso_full_data <- results_per_fold[results_per_fold$type == "loso" & results_per_fold$label_set == "Full subtypes", ]

loso_full_data %>% filter(model == "SVM") %>% mutate(type = ifelse(fold %in% pediatric_cohorts, "Paediatric", "Adult")) %>% group_by(type) %>% summarise(mean_kappa = mean(kappa))

loso_full_prep <- prepare_fold_data(loso_full_data) 
pc <- ggplot(loso_full_prep$data, aes(x = fold, y = kappa, color = model, group = model)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, colour = loso_full_prep$label_colors),
        legend.position = "none") +
  labs(title = "LOSO: Full subtypes", x = "Cohort (Fold)", y = "Cohen's Kappa", color = "Model") + 
  scale_color_manual(values = colours, breaks = rev(names(colours)), limits = rev(names(colours))) + 
  ylim(0.75, 1)

# LOSO: Collapsed classes - order by mean kappa
loso_collapsed_data <- results_per_fold[results_per_fold$type == "loso" & results_per_fold$label_set == "Collapsed classes", ]

loso_collapsed_data %>% filter(model == "SVM") %>% mutate(type = ifelse(fold %in% pediatric_cohorts, "Paediatric", "Adult")) %>% group_by(type) %>% summarise(mean_kappa = mean(kappa))

loso_collapsed_prep <- prepare_fold_data(loso_collapsed_data)
pd <- ggplot(loso_collapsed_prep$data, aes(x = fold, y = kappa, color = model, group = model)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, colour = loso_collapsed_prep$label_colors),
        legend.position = "right") +
  labs(title = "LOSO: Collapsed classes", x = "Cohort (Fold)", y = "Cohen's Kappa", color = "Model") + 
  scale_color_manual(values = colours, breaks = rev(names(colours)), limits = rev(names(colours))) + 
  ylim(0.75, 1)

# Combine plots with merged legends
library(patchwork)
library(cowplot)
library(grid)

# Dummy plot for pediatric/adult legend
ped_adult_legend_data <- data.frame(cohort_type = c("Pediatric", "Adult"), x = 1:2, y = 1:2)
ped_adult_legend_plot <- ggplot(ped_adult_legend_data, aes(x = x, y = y, fill = cohort_type)) +
  geom_bar(stat = "identity", width = 0.5) +
  scale_fill_manual(values = c("Adult" = "#7eb0d5", "Pediatric" = "#fd7f6f"), name = "Cohort Type") +
  theme_void() +
  theme(legend.position = "right", legend.title = element_text(size = 10),
        legend.text = element_text(size = 9), legend.key = element_rect(fill = "white", color = NA),
        plot.margin = margin(0, 0, 0, 0)) +
  xlim(0, 0) + ylim(0, 0)

# Combine in 2x2 grid with legends
plots_grid <- (pa + theme(legend.position = "none") + pb + theme(legend.position = "none")) / 
              (pc + theme(legend.position = "none") + pd + theme(legend.position = "none"))
model_legend <- get_legend(pa + theme(legend.position = "right"))
ped_adult_legend <- get_legend(ped_adult_legend_plot)
combined_legends <- plot_grid(model_legend, ped_adult_legend, ncol = 1, align = "v")
final_plot <- plot_grid(plots_grid, combined_legends, ncol = 2, rel_widths = c(4, 0.6))

print(final_plot)
```

## -- Cohort-Level Performance

```{r}
library(ggplot2)
library(dplyr)

# Extract LOSO performance per cohort
loso_performance <- outer_merged$detailed_performance$loso$OvR_Ensemble

cohort_performance <- data.frame(
  cohort = names(loso_performance),
  kappa = sapply(loso_performance, function(x) x$kappa),
  accuracy = sapply(loso_performance, function(x) x$accuracy),
  n = sapply(loso_performance, function(x) length(x$predictions)),
  stringsAsFactors = FALSE
)

# Cohort subtype distribution
cohort_subtype_dist <- data.frame(
  cohort = filtered_study_names,
  subtype = filtered_leukemia_subtypes,
  stringsAsFactors = FALSE
)
cohort_subtype_dist$subtype_merged <- modify_classes(cohort_subtype_dist$subtype)
cohort_subtype_dist$subtype_display <- fix_names(cohort_subtype_dist$subtype_merged)

# Merge and add pediatric/adult classification
cohort_analysis <- cohort_performance
cohort_analysis$cohort_type <- ifelse(
  cohort_analysis$cohort %in% c("AAML0531", "AAML1031", "AAML03P1"), "Pediatric", "Adult"
)

print(cohort_analysis)
cohort_analysis %>% group_by(cohort_type) %>% summarise(mean = mean(kappa), sd = sd(kappa))
```

```{r}
# Boxplot: pediatric vs adult performance
cohort_analysis_long <- cohort_analysis %>%
  tidyr::pivot_longer(cols = c(kappa, accuracy), names_to = "metric", values_to = "value")

p_ped_adult <- ggplot(cohort_analysis_long, aes(x = cohort_type, y = value, fill = cohort_type)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.2, size = 3, alpha = 0.6) +
  facet_wrap(~metric, scales = "free_y", 
             labeller = labeller(metric = c(kappa = "Cohen's Kappa", accuracy = "Accuracy"))) +
  scale_fill_manual(values = c("Adult" = "#7eb0d5", "Pediatric" = "#fd7f6f")) +
  theme_bw() +
  labs(title = "LOSO Performance: Pediatric vs Adult Cohorts", x = "Cohort Type", 
       y = "Performance", fill = "Cohort Type") +
  theme(legend.position = "none")

print(p_ped_adult)
ggsave("../writing/figures_new/pediatric_vs_adult_performance.png", p_ped_adult, width = 8, height = 5, dpi = 300)
```

## -- Data Preparation: Sample-Level Analysis

```{r}
# Build sample-level dataframe with LOSO predictions and metadata
loso_predictions_list <- list()
loso_truth_list <- list()
loso_fold_list <- list()
loso_indices_list <- list()

for (fold in names(outer_merged$detailed_performance$loso$svm)) {
  fold_prob_matrix <- outer_merged$ensemble_matrices$loso$ovr_ensemble[[fold]]
  predictions <- apply(fold_prob_matrix[,!colnames(fold_prob_matrix) %in% c("y", "outer_fold", "sample_indices")], 1, which.max)
  predictions <- colnames(fold_prob_matrix)[!colnames(fold_prob_matrix) %in% c("y", "outer_fold", "sample_indices")][predictions]
  truth <- as.character(fold_prob_matrix$y)
  sample_indices <- fold_prob_matrix$sample_indices
  
  loso_predictions_list[[fold]] <- predictions
  loso_truth_list[[fold]] <- truth
  loso_fold_list[[fold]] <- rep(fold, length(predictions))
  loso_indices_list[[fold]] <- sample_indices
}

# Combine predictions
sample_level_df <- data.frame(
  prediction = unlist(loso_predictions_list),
  truth = unlist(loso_truth_list),
  cohort = unlist(loso_fold_list),
  sample_index = unlist(loso_indices_list),
  stringsAsFactors = FALSE
)
sample_level_df$correct <- sample_level_df$prediction == sample_level_df$truth

# Add age and apply class transformations
sample_level_df$age_days <- meta$age_days[sample_level_df$sample_index + 1]
sample_level_df$age_years <- sample_level_df$age_days / 365.25
sample_level_df$truth_merged <- modify_classes(sample_level_df$truth)
sample_level_df$truth_display <- fix_names(sample_level_df$truth_merged)
sample_level_df$prediction_merged <- modify_classes(sample_level_df$prediction)
sample_level_df$prediction_display <- fix_names(sample_level_df$prediction_merged)
sample_level_df$is_pediatric <- sample_level_df$cohort %in% c("AAML0531", "AAML1031", "AAML03P1")

# Define rare subtypes
subtype_counts <- table(sample_level_df$truth_display)
rare_subtypes <- names(subtype_counts[subtype_counts < 50])
sample_level_df$is_rare <- sample_level_df$truth_display %in% rare_subtypes

# Alias for downstream analyses
sample_level_age <- sample_level_df
```

## -- Visualization: Age Effect on Performance

```{r}
# Binned kappa by age
age_bins <- seq(0, 90, by = 5)
age_performance <- data.frame()

for (i in 1:(length(age_bins)-1)) {
  age_min <- age_bins[i]
  age_max <- age_bins[i+1]
  subset_data <- sample_level_age[sample_level_age$age_years >= age_min & 
                                   sample_level_age$age_years < age_max, ]
  
  if (nrow(subset_data) >= 10) {
    cm <- caret::confusionMatrix(
      factor(subset_data$prediction_display, levels = unique(c(subset_data$truth_display, subset_data$prediction_display))),
      factor(subset_data$truth_display, levels = unique(c(subset_data$truth_display, subset_data$prediction_display)))
    )
    
    age_performance <- rbind(age_performance, data.frame(
      age_mid = (age_min + age_max) / 2,
      age_label = sprintf("%d-%d", age_min, age_max),
      n = nrow(subset_data),
      accuracy = mean(subset_data$correct),
      kappa = as.numeric(cm$overall["Kappa"]),
      se_accuracy = sd(subset_data$correct) / sqrt(nrow(subset_data))
    ))
  }
}

p_age_kappa <- ggplot(age_performance, aes(x = age_mid, y = kappa)) +
  geom_point(aes(size = n), alpha = 0.7, color = "#7eb0d5") +
  geom_smooth(method = "loess", se = TRUE, color = "#7eb0d5", fill = "#7eb0d5", 
              alpha = 0.2, formula = y ~ x) +
  geom_vline(xintercept = 18, linetype = "dashed", color = "gray40", alpha = 0.5) +
  annotate("text", x = 18, y = min(age_performance$kappa) - 0.02, 
           label = "Pediatric/Adult\nBoundary", vjust = 1, size = 3, color = "gray40") +
  scale_size_continuous(range = c(3, 12), name = "Sample Size") +
  theme_bw() +
  labs(title = "Prediction Performance by Patient Age",
       subtitle = sprintf("Based on LOSO validation (n=%d patients)", nrow(sample_level_age)),
       x = "Age (years)", y = "Cohen's Kappa") +
  ylim(min(age_performance$kappa) - 0.15, 1) +
  theme(text = element_text(size = 12))

print(p_age_kappa)
ggsave("../writing/figures_new/age_effect_on_performance.png", p_age_kappa, width = 10, height = 6, dpi = 300)
```

## -- Per-Subtype F1 by Age Analysis

```{r}
# Calculate F1 per subtype per 15-year age bin
age_bins_5yr <- seq(0, 90, by = 20)
subtypes_unique <- unique(sample_level_age$truth_display)
f1_by_age_subtype <- data.frame()

for (i in 1:(length(age_bins_5yr)-1)) {
  age_min <- age_bins_5yr[i]
  age_max <- age_bins_5yr[i+1]
  age_mid <- (age_min + age_max) / 2
  age_label <- sprintf("%d-%d", age_min, age_max)
  
  age_bin_data <- sample_level_age[sample_level_age$age_years >= age_min & 
                                     sample_level_age$age_years < age_max, ]
  if (nrow(age_bin_data) == 0) next
  
  for (subtype_i in subtypes_unique) {
    tp <- sum(age_bin_data$truth_display == subtype_i & age_bin_data$prediction_display == subtype_i)
    fp <- sum(age_bin_data$truth_display != subtype_i & age_bin_data$prediction_display == subtype_i)
    fn <- sum(age_bin_data$truth_display == subtype_i & age_bin_data$prediction_display != subtype_i)
    n_true <- sum(age_bin_data$truth_display == subtype_i)
    
    if (n_true >= 5) {
      precision <- ifelse(tp + fp > 0, tp / (tp + fp), NA)
      recall <- ifelse(tp + fn > 0, tp / (tp + fn), NA)
      f1_score <- ifelse(!is.na(precision) & !is.na(recall) & (precision + recall) > 0,
                         2 * precision * recall / (precision + recall), NA)
      
      f1_by_age_subtype <- rbind(f1_by_age_subtype, data.frame(
        age_bin = age_label, age_min = age_min, age_max = age_max, age_mid = age_mid,
        truth_display = subtype_i, n = n_true, tp = tp, fp = fp, fn = fn,
        precision = precision, recall = recall, f1_score = f1_score
      ))
    }
  }
}

# Filter to subtypes with data across multiple age ranges
subtypes_multi_age <- f1_by_age_subtype %>%
  group_by(truth_display) %>%
  summarise(n_age_bins = n(), .groups = "drop") %>%
  filter(n_age_bins >= 2)

f1_by_age_plot <- f1_by_age_subtype %>%
  filter(truth_display %in% subtypes_multi_age$truth_display)

# Faceted plot
p_f1_age_subtype_facet <- ggplot(f1_by_age_plot, aes(x = age_mid, y = f1_score)) +
  geom_line(linewidth = 1, color = "#7eb0d5", alpha = 0.8) +
  geom_point(aes(size = n), color = "#7eb0d5", alpha = 0.7) +
  geom_vline(xintercept = 18, linetype = "dashed", color = "gray40", alpha = 0.4, linewidth = 0.5) +
  facet_wrap(~truth_display, ncol = 4, scales = "free_y") +
  scale_size_continuous(range = c(2, 6), name = "Sample Size (n)",
                        breaks = c(10, 25, 50), labels = c("n=10", "n=25", "n=50")) +
  theme_bw() +
  labs(title = "Per-Subtype F1 Score by Patient Age",
       subtitle = "Dashed line marks pediatric/adult boundary (18 years)",
       x = "Age (years)", y = "F1 Score") +
  theme(strip.background = element_rect(fill = "gray95"),
        strip.text = element_text(size = 9, face = "bold"),
        panel.grid.minor = element_blank(),
        legend.position = "bottom") +
  ylim(0, 1)

print(p_f1_age_subtype_facet)
ggsave("../writing/figures_new/f1_by_age_and_subtype_faceted.png", p_f1_age_subtype_facet, width = 14, height = 10, dpi = 300)
```

## -- Data Preparation: Subtype Enrichment Analysis

```{r}
# Get F1 scores per subtype from per-class summaries
per_class_f1 <- outer_merged$per_class_summaries$cv %>%
  filter(Model == "OvR_Ensemble") %>%
  select(Class, Mean_F1)
per_class_f1$Class <- gsub("Class: ", "", per_class_f1$Class)
per_class_f1$Class <- fix_names(per_class_f1$Class)

# Analyze subtype distribution in pediatric vs adult samples
subtype_enrichment <- sample_level_age %>%
  group_by(truth_display, is_pediatric) %>%
  summarise(n = n(), accuracy = mean(correct), .groups = "drop") %>%
  tidyr::pivot_wider(names_from = is_pediatric, values_from = c(n, accuracy),
                     values_fill = list(n = 0, accuracy = NA)) %>%
  rename(n_adult = `n_FALSE`, n_pediatric = `n_TRUE`,
         accuracy_adult = `accuracy_FALSE`, accuracy_pediatric = `accuracy_TRUE`)

# Merge with F1 scores and calculate enrichment
subtype_enrichment <- merge(subtype_enrichment, per_class_f1, 
                            by.x = "truth_display", by.y = "Class", all.x = TRUE)

total_adult <- sum(subtype_enrichment$n_adult)
total_pediatric <- sum(subtype_enrichment$n_pediatric)
subtype_enrichment$prop_adult <- subtype_enrichment$n_adult / total_adult
subtype_enrichment$prop_pediatric <- subtype_enrichment$n_pediatric / total_pediatric
subtype_enrichment$enrichment_ratio <- (subtype_enrichment$prop_pediatric + 0.001) / 
                                       (subtype_enrichment$prop_adult + 0.001)
subtype_enrichment$overall_accuracy <- 
  (subtype_enrichment$n_adult * subtype_enrichment$accuracy_adult + 
   subtype_enrichment$n_pediatric * subtype_enrichment$accuracy_pediatric) /
  (subtype_enrichment$n_adult + subtype_enrichment$n_pediatric)

# Chi-square/Fisher tests for enrichment classification
subtype_enrichment$chi_sq_p <- NA
for (i in 1:nrow(subtype_enrichment)) {
  cont_table <- matrix(c(subtype_enrichment$n_pediatric[i], 
                         total_pediatric - subtype_enrichment$n_pediatric[i],
                         subtype_enrichment$n_adult[i], 
                         total_adult - subtype_enrichment$n_adult[i]),
                       nrow = 2, byrow = TRUE)
  test_result <- if (min(cont_table) < 5) fisher.test(cont_table) else chisq.test(cont_table)
  subtype_enrichment$chi_sq_p[i] <- test_result$p.value
}
subtype_enrichment$chi_sq_p_adj <- p.adjust(subtype_enrichment$chi_sq_p, method = "BH")
subtype_enrichment$enrichment_class <- ifelse(
  subtype_enrichment$chi_sq_p_adj < 0.05,
  ifelse(subtype_enrichment$enrichment_ratio > 1, "Pediatric-enriched",
         ifelse(subtype_enrichment$enrichment_ratio < 1, "Adult-enriched", "Similar")),
  "Similar"
)

# Label samples by enrichment class
sample_level_age$subtype_enrichment <- subtype_enrichment$enrichment_class[
  match(sample_level_age$truth_display, subtype_enrichment$truth_display)
]
```

## -- Data Preparation: Per-Class F1 by Cohort Type

```{r}
# Function to calculate F1 for a specific subtype and cohort subset
calc_f1_for_subset <- function(data_subset, subtype_name, all_data) {
  if (nrow(data_subset) < 5) return(NULL)
  
  n_total <- nrow(data_subset)
  n_correct <- sum(data_subset$prediction_display == data_subset$truth_display)
  sensitivity <- n_correct / n_total
  
  n_predicted_as_this <- sum(all_data$prediction_display == subtype_name)
  if (n_predicted_as_this > 0) {
    n_correct_total <- sum(all_data$prediction_display == subtype_name & 
                           all_data$truth_display == subtype_name)
    precision <- n_correct_total / n_predicted_as_this
    f1 <- if (precision + sensitivity > 0) 2 * (precision * sensitivity) / (precision + sensitivity) else 0
  } else {
    precision <- 0
    f1 <- 0
  }
  
  list(n = n_total, sensitivity = sensitivity, precision = precision, f1 = f1, n_correct = n_correct)
}

# Calculate F1 per class per cohort type
f1_per_class_cohort_list <- list()

for (subtype in unique(sample_level_age$truth_display)) {
  subset_all <- sample_level_age[sample_level_age$truth_display == subtype, ]
  if (nrow(subset_all) < 10) next
  
  subset_ped <- subset_all[subset_all$is_pediatric == TRUE, ]
  subset_adult <- subset_all[subset_all$is_pediatric == FALSE, ]
  all_ped <- sample_level_age[sample_level_age$is_pediatric == TRUE, ]
  all_adult <- sample_level_age[sample_level_age$is_pediatric == FALSE, ]
  
  if (nrow(subset_ped) >= 5) {
    metrics_ped <- calc_f1_for_subset(subset_ped, subtype, all_ped)
    if (!is.null(metrics_ped)) {
      f1_per_class_cohort_list[[length(f1_per_class_cohort_list) + 1]] <- data.frame(
        subtype = subtype, is_pediatric = TRUE, cohort_type = "Pediatric",
        n = metrics_ped$n, sensitivity = metrics_ped$sensitivity,
        precision = metrics_ped$precision, f1 = metrics_ped$f1, n_correct = metrics_ped$n_correct
      )
    }
  }
  
  if (nrow(subset_adult) >= 5) {
    metrics_adult <- calc_f1_for_subset(subset_adult, subtype, all_adult)
    if (!is.null(metrics_adult)) {
      f1_per_class_cohort_list[[length(f1_per_class_cohort_list) + 1]] <- data.frame(
        subtype = subtype, is_pediatric = FALSE, cohort_type = "Adult",
        n = metrics_adult$n, sensitivity = metrics_adult$sensitivity,
        precision = metrics_adult$precision, f1 = metrics_adult$f1, n_correct = metrics_adult$n_correct
      )
    }
  }
}

f1_per_class_cohort <- do.call(rbind, f1_per_class_cohort_list)
f1_per_class_cohort <- merge(f1_per_class_cohort,
  subtype_enrichment[, c("truth_display", "enrichment_class", "n_pediatric", "n_adult")],
  by.x = "subtype", by.y = "truth_display", all.x = TRUE)

# Filter to enriched classes and sort by mean F1
f1_per_class_enriched <- f1_per_class_cohort[
  f1_per_class_cohort$enrichment_class %in% c("Adult-enriched", "Pediatric-enriched"), ]

f1_avg <- f1_per_class_enriched %>%
  group_by(subtype) %>%
  summarise(avg_f1 = mean(f1, na.rm = TRUE), .groups = "drop")

f1_per_class_enriched <- merge(f1_per_class_enriched, f1_avg, by = "subtype")
f1_per_class_enriched <- f1_per_class_enriched[order(f1_per_class_enriched$avg_f1), ]
f1_per_class_enriched$subtype <- factor(f1_per_class_enriched$subtype, 
                                        levels = unique(f1_per_class_enriched$subtype))
```

## -- Visualization: Per-Class F1 Score by Cohort Type and Enrichment Status

```{r}
library(patchwork)

# Split by enrichment class
dat_top <- f1_per_class_enriched %>% filter(enrichment_class == "Adult-enriched")
dat_bottom <- f1_per_class_enriched %>% filter(enrichment_class != "Adult-enriched")

p_top <- ggplot(dat_top, aes(x = subtype, y = f1, fill = cohort_type)) +
  geom_col(position = position_dodge(width = 0.85), width = 0.75, alpha = 0.85) +
  geom_text(aes(label = sprintf("%.2f", f1)), position = position_dodge(width = 0.85),
            hjust = 1.05, size = 2.7, color = "black") +
  coord_flip() +
  scale_fill_manual(values = c("Pediatric" = "#e67e62", "Adult" = "#5a9acb"), name = "Cohort type") +
  scale_y_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.05))) +
  labs(title = "LOSO performance per AML subtype", subtitle = "Adult enriched classes",
       x = "Subtype", y = "F1 score") +
  theme_bw() +
  theme(legend.position = "none", axis.text.y = element_text(size = 8),
        panel.grid.major.y = element_blank())

p_bottom <- ggplot(dat_bottom, aes(x = subtype, y = f1, fill = cohort_type)) +
  geom_col(position = position_dodge(width = 0.85), width = 0.75, alpha = 0.85) +
  geom_text(aes(label = sprintf("%.2f", f1)), position = position_dodge(width = 0.85),
            hjust = 1.05, size = 2.7, color = "black") +
  coord_flip() +
  scale_fill_manual(values = c("Pediatric" = "#e67e62", "Adult" = "#5a9acb"), name = "Cohort type") +
  scale_y_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.05))) +
  labs(subtitle = "Pediatric enriched classes", x = "Subtype", y = "F1 score") +
  theme_bw() +
  theme(legend.position = "bottom", axis.text.y = element_text(size = 8),
        panel.grid.major.y = element_blank())

p_f1_per_class <- p_top / p_bottom + plot_layout(heights = c(3, 8))
print(p_f1_per_class)
ggsave("../writing/figures_new/figures_class_perf.png", p_f1_per_class, width = 10, height = 10, dpi = 300)
```

## -- Visualization: Subtype Enrichment and Performance

```{r}
p_enrichment <- ggplot(subtype_enrichment, 
                       aes(x = log2(enrichment_ratio), y = Mean_F1, 
                           color = enrichment_class, size = n_pediatric + n_adult)) +
  geom_point(alpha = 0.7) +
  geom_text(aes(label = truth_display), hjust = -0.1, vjust = 0.5, size = 3, show.legend = FALSE) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.5) +
  scale_color_manual(values = c("Pediatric-enriched" = "#fd7f6f", 
                                "Adult-enriched" = "#7eb0d5", "Similar" = "gray60"),
                     name = "Enrichment") +
  scale_size_continuous(range = c(2, 10), name = "Total Samples") +
  theme_bw() +
  labs(title = "Subtype Performance (F1) vs Pediatric/Adult Enrichment",
       subtitle = "Negative log2 ratio = Adult-enriched, Positive = Pediatric-enriched",
       x = "Log2(Pediatric/Adult Enrichment Ratio)", y = "F1 Score") +
  theme(legend.position = "bottom") #+
  #xlim(min(log2(subtype_enrichment$enrichment_ratio)) / 1.3,
  #max(log2(subtype_enrichment$enrichment_ratio)) * 1.3) +
  #ylim(0.7, 1)

print(p_enrichment)
ggsave("../writing/figures_new/subtype_enrichment_performance.png", p_enrichment, width = 10, height = 8, dpi = 300)
```


## -- Visualization: F1 Scores per LOSO Fold (Full vs Merged)

```{r}
# Minimum number of samples per subtype in a fold to show that fold's F1 (avoids unstable F1 from tiny n)
min_n_per_subtype_in_fold <- 5L

# Extract per-fold F1 scores for both full and merged classes
# Helper function to get per-class F1 per fold for LOSO (includes n = count of subtype in fold)
get_f1_per_fold_loso <- function(res_obj, label_set_name) {
  folds <- names(res_obj$detailed_performance$loso$svm)
  f1_list <- list()
  
  for (fold in folds) {
    # Get predictions and truth for this fold
    fold_perf <- res_obj$detailed_performance$loso$OvR_Ensemble[[fold]]
    predictions <- fold_perf$predictions
    truth <- fold_perf$truth
    
    # Calculate F1 per class for this fold
    classes <- unique(c(as.character(predictions), as.character(truth)))
    
    for (cls in classes) {
      tp <- sum(truth == cls & predictions == cls)
      fp <- sum(truth != cls & predictions == cls)
      fn <- sum(truth == cls & predictions != cls)
      n_in_fold <- sum(truth == cls)  # number of samples of this subtype in this fold
      
      precision <- ifelse(tp + fp > 0, tp / (tp + fp), 0)
      recall <- ifelse(tp + fn > 0, tp / (tp + fn), 0)
      f1 <- ifelse(precision + recall > 0, 2 * precision * recall / (precision + recall), 0)
      
      f1_list[[length(f1_list) + 1]] <- data.frame(
        fold = fold,
        class = cls,
        f1 = f1,
        n_in_fold = n_in_fold,
        label_set = label_set_name,
        stringsAsFactors = FALSE
      )
    }
  }
  
  do.call(rbind, f1_list)
}

# Get F1 per fold for both full and merged
f1_full_loso <- get_f1_per_fold_loso(outer_unmerged, "Full subtypes")
f1_merged_loso <- get_f1_per_fold_loso(outer_merged, "Collapsed classes")

# Clean class names
f1_full_loso$class <- fix_names(f1_full_loso$class)
f1_merged_loso$class <- fix_names(f1_merged_loso$class)

# Combine datasets
f1_loso_combined <- rbind(f1_full_loso, f1_merged_loso)
f1_loso_combined$label_set <- factor(f1_loso_combined$label_set, 
                                     levels = c("Full subtypes", "Collapsed classes"))

# Keep only (fold, class) rows where subtype has at least min_n samples in that fold
f1_loso_combined <- f1_loso_combined %>% filter(n_in_fold >= min_n_per_subtype_in_fold)

# Order classes by mean F1 (using merged as reference)
# class_order <- f1_merged_loso %>%
#   group_by(class) %>%
#   summarise(mean_f1 = mean(f1, na.rm = TRUE), .groups = "drop") %>%
#   arrange(mean_f1) %>%
#   pull(class)

# Keep only classes present in both datasets for comparison
#common_classes <- intersect(unique(f1_full_loso$class), unique(f1_merged_loso$class))
f1_loso_plot <- f1_loso_combined # %>%
#  filter(class %in% common_classes)

f1_loso_plot$class <- factor(f1_loso_plot$class, 
                             levels = rev(order_AML))#[class_order %in% common_classes])
f1_loso_plot$label_set <- NULL
f1_loso_plot <- unique(f1_loso_plot)
# Create faceted plot with dots per LOSO fold
p_f1_loso_facet <- ggplot(f1_loso_plot, aes(x = f1, y = class, color = fold)) +
  geom_point(size = 2.5, alpha = 0.7) +
  geom_vline(xintercept = 0.8, linetype = "dashed", color = "gray50", alpha = 0.5) +
  #facet_wrap(~ label_set, ncol = 2) +
  # scale_color_manual(
  #   values = c("AAML0531" = "#fd7f6f", "AAML1031" = "#e67e62", "AAML03P1" = "#d66b5b",
  #              "BEAT_AML_Waves_1and2" = "#7eb0d5", "TCGA_LAML" = "#5a9acb", 
  #              "LEUCEGENE" = "#4a8ab8", "TARGET_AML" = "#3a7aa8"),
  #   name = "LOSO Fold"
  # ) +
  theme_bw() +
  labs(
    title = "F1 Scores per AML Subtype",
    subtitle = sprintf("Each dot = one LOSO fold (only folds with n >= %d samples per subtype)", min_n_per_subtype_in_fold),
    x = "F1 Score",
    y = "AML Subtype"
  ) +
  theme(
    strip.background = element_rect(fill = "gray95"),
    strip.text = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 7),
    legend.position = "bottom",
    panel.grid.major.y = element_line(color = "gray90", linewidth = 0.3)
  ) +
  xlim(0, 1)

print(p_f1_loso_facet)
ggsave("../writing/figures_new/f1_per_loso_fold_full_vs_merged.png", p_f1_loso_facet, 
       width = 14, height = 12, dpi = 300)
```

## -- Visualization: Delta F1 from Class Collapsing

```{r}
# Only include folds where the collapsed class has at least this many samples (avoids unstable delta from tiny n)
min_n_delta_f1 <- 5L

# Calculate delta F1 (F1_collapsed - F1_full) for collapsed class groups
# Focus on MDS-related, other KMT2A, and MECOM rearrangement classes that were merged

# Define the merge mapping: which full classes map to which collapsed class
merge_mapping <- list(
  "MDS-related" = c("MDS-related, gene mutations", "MDS-related, cytogenetic abnormalities", 
                    "MDS-related, TP53 mutated", "MDS.r"),
  "Other KMT2A rearrangements" = c("KMT2A::ELL", "KMT2A::MLLT10","KMT2A::MLLT1", "KMT2A::MLLT4", "KMT2A::SEPT6"),
  "MECOM rearrangement" = c("GATA2;MECOM", "MECOM, other")
)

# Calculate mean F1 per fold for full subclasses and collapsed classes (only folds with n >= min_n_delta_f1)
delta_f1_list <- list()

for (collapsed_class in names(merge_mapping)) {
  full_classes <- merge_mapping[[collapsed_class]]
  
  # Get F1 for the collapsed class per fold (from merged results); keep only folds with n >= min_n
  collapsed_f1 <- f1_merged_loso %>%
    filter(class == collapsed_class, n_in_fold >= min_n_delta_f1) %>%
    select(fold, f1_collapsed = f1)
  
  # Get mean F1 across the constituent full classes per fold
  # (weighted average would be better but mean is simpler for visualization)
  full_f1 <- f1_full_loso %>%
    filter(class %in% full_classes) %>%
    group_by(fold) %>%
    summarise(f1_full = mean(f1, na.rm = TRUE), .groups = "drop")
  
  # Calculate delta
  delta_df <- merge(collapsed_f1, full_f1, by = "fold")
  delta_df$delta_f1 <- delta_df$f1_collapsed - delta_df$f1_full
  delta_df$group <- collapsed_class
  
  delta_f1_list[[collapsed_class]] <- delta_df
}

delta_f1_data <- do.call(rbind, delta_f1_list)

# Add cohort type for coloring
delta_f1_data$cohort_type <- ifelse(
  delta_f1_data$fold %in% c("AAML0531", "AAML1031", "AAML03P1"), 
  "Pediatric", "Adult"
)

# Calculate summary statistics for reporting
delta_summary <- delta_f1_data %>%
  group_by(group) %>%
  summarise(
    mean_delta = mean(delta_f1, na.rm = TRUE),
    sd_delta = sd(delta_f1, na.rm = TRUE),
    min_f1_full = min(f1_full, na.rm = TRUE),
    max_f1_full = max(f1_full, na.rm = TRUE),
    min_f1_collapsed = min(f1_collapsed, na.rm = TRUE),
    max_f1_collapsed = max(f1_collapsed, na.rm = TRUE),
    .groups = "drop"
  )

cat("Delta F1 Summary (Collapsed - Full):\n")
cat("=====================================\n")
print(delta_summary)

# Bar plot with individual fold points showing delta F1
p_delta_f1 <- ggplot(delta_f1_data, aes(x = group, y = delta_f1)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_boxplot(aes(fill = group), alpha = 0.3, outlier.shape = NA, width = 0.5) +
  geom_point(aes(color = cohort_type, shape = cohort_type), 
             size = 3, alpha = 0.8, position = position_jitter(width = 0.1, seed = 42)) +
  scale_fill_manual(values = c("MDS-related" = "#d62728", 
                               "Other KMT2A rearrangements" = "#ff7f0e",
                               "MECOM rearrangement" = "#17becf"),
                    guide = "none") +
  scale_color_manual(values = c("Adult" = "#7eb0d5", "Pediatric" = "#fd7f6f"),
                     name = "Cohort Type") +
  scale_shape_manual(values = c("Adult" = 16, "Pediatric" = 17),
                     name = "Cohort Type") +
  theme_bw() +
  labs(
    title = "F1 Score Improvement from Class Collapsing (LOSO)",
    subtitle = paste(
      sprintf("Folds with n >= %d per collapsed class. ", min_n_delta_f1),
      sprintf("MDS-related: mean dF1 = +%.2f", delta_summary$mean_delta[delta_summary$group == "MDS-related"]),
      sprintf("| Other KMT2A: +%.2f", delta_summary$mean_delta[delta_summary$group == "Other KMT2A rearrangements"]),
      if ("MECOM rearrangement" %in% delta_summary$group)
        sprintf("| MECOM: +%.2f", delta_summary$mean_delta[delta_summary$group == "MECOM rearrangement"]) else "",
      sep = " "
    ),
    x = "Collapsed Class Group",
    y = expression(Delta*"F1 (Collapsed - Full)")
  ) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(size = 10)
  ) +
  ylim(-0.1, max(delta_f1_data$delta_f1) + 0.05)

print(p_delta_f1)
ggsave("../writing/figures_new/delta_f1_collapsed_classes.png", p_delta_f1, 
       width = 8, height = 6, dpi = 300)

# Print detailed results for manuscript text
cat("\n\nFor manuscript text:\n")
cat("====================\n")
for (grp in unique(delta_f1_data$group)) {
  grp_data <- delta_f1_data[delta_f1_data$group == grp, ]
  grp_summary <- delta_summary[delta_summary$group == grp, ]
  cat(sprintf("\n%s:\n", grp))
  cat(sprintf("  Full F1 range: %.2f - %.2f\n", grp_summary$min_f1_full, grp_summary$max_f1_full))
  cat(sprintf("  Collapsed F1 range: %.2f - %.2f\n", grp_summary$min_f1_collapsed, grp_summary$max_f1_collapsed))
  cat(sprintf("  Mean dF1: +%.2f (SD: %.2f)\n", grp_summary$mean_delta, grp_summary$sd_delta))
}
```

# Gene selection analysis

```{r}
counts <- data.table::fread("../data/counts_20aug25.csv", drop = 1)
counts <- data.frame(data.table::transpose(counts))
counts_filtered <- counts[filter,]

library(reticulate)
use_virtualenv("../.venv/", required = TRUE)

# Ensure our local `python` folder (with `transformers.py`) is on sys.path
py_run_string("
import sys, os
local_python = os.path.abspath('../python')
if local_python not in sys.path:
    sys.path.insert(0, local_python)
")

# Load full fitted pipeline (joblib). Steps: DESeq2 norm -> MVG selection -> scaling.
py_run_string("import joblib")
py_run_string("pipeline = joblib.load('../data/out/final_models/pipelines/pipeline_ngenes_2000.pkl')")
pipeline_2000 <- py$pipeline

# Get full gene name vector (for all genes)
counts_gene_names <- data.table::fread("../data/counts_20aug25.csv", select = 1, header = TRUE)[[1]]

# Indices of selected MVGs from the pipeline (0-based in Python, 1-based in R)
mvgs_1based <- py_to_r(py_eval("list(pipeline.named_steps['feature_selection'].mvgs_)"))
mvgs_1based <- as.integer(mvgs_1based) + 1L

# Build matrix of filtered counts (samples x all genes)
X_counts <- as.matrix(counts_filtered)
mode(X_counts) <- "numeric"
py$X <- X_counts

# DESeq2-only normalisation on ALL genes (no MVG filter, no scaler)
counts_deseq2_all <- py_to_r(pipeline_2000$named_steps$DEseq2$transform(py$X))
rownames(counts_deseq2_all) <- rownames(counts_filtered)
colnames(counts_deseq2_all) <- counts_gene_names
```

## Eta-squared: study vs subtype association per gene

```{r}
# Vectorized eta-squared (proportion of variance explained) per gene for study and subtype.
X <- as.matrix(counts_deseq2_all)
n <- nrow(X)
mean_g <- colMeans(X)
SS_total <- colSums((X - rep(mean_g, each = n))^2)
SS_total[SS_total <= 0] <- NA

study_f <- factor(filtered_study_names)
rs_study <- rowsum(X, study_f)
n_study <- as.vector(table(study_f))
mean_by_study <- rs_study / n_study
SS_between_study <- colSums(n_study * (mean_by_study - rep(mean_g, each = nrow(mean_by_study)))^2)
eta2_study <- SS_between_study / SS_total

subtype_f <- factor(modify_classes(filtered_leukemia_subtypes))
rs_subtype <- rowsum(X, subtype_f)
n_subtype <- as.vector(table(subtype_f))
mean_by_subtype <- rs_subtype / n_subtype
SS_between_subtype <- colSums(n_subtype * (mean_by_subtype - rep(mean_g, each = nrow(mean_by_subtype)))^2)
eta2_subtype <- SS_between_subtype / SS_total

eta2_study[is.nan(eta2_study)] <- NA
eta2_subtype[is.nan(eta2_subtype)] <- NA

df_eta2 <- data.frame(
  gene = counts_gene_names,
  eta2_study = as.numeric(eta2_study),
  eta2_subtype = as.numeric(eta2_subtype),
  selected = seq_along(counts_gene_names) %in% mvgs_1based,
  stringsAsFactors = FALSE
)
```

```{r}
library(ggplot2)
p_eta2 <- ggplot(df_eta2, aes(x = eta2_study, y = eta2_subtype)) +
  geom_point(data = subset(df_eta2, !selected), color = "black", alpha = 0.35, size = 0.6) +
  geom_point(data = subset(df_eta2, selected), color = "red", alpha = 0.6, size = 0.8) +
  theme_bw() +
  labs(
    x = expression(eta^2 ~ "(study | gene expression)"),
    y = expression(eta^2 ~ "(subtype | gene expression)"),
    title = "Study vs subtype association per gene"
  )
p_eta2
#ggsave("../writing/figures_new/eta2_study_vs_subtype.png", p_eta2, width = 6, height = 5, dpi = 300)
```

## t-SNE on MVGs (PCA 50)

```{r}
library(Rtsne)

# MVG expression (DESeq2-normalised), then scale for PCA
X_mvg <- scale(counts_deseq2_all[, mvgs_1based])
# PCA on MVGs, then use first 50 PCs for fast t-SNE
pca_mvg <- prcomp(X_mvg, scale. = FALSE)
pcs_50 <- pca_mvg$x[, 1:50]
# t-SNE on 50 PC scores
set.seed(42)
tsne_mvg <- Rtsne(
  pcs_50,
  dims = 2,
  perplexity = min(30, (nrow(pca_mvg$x) - 1) %/% 3),
  max_iter = 1000,
  pca = FALSE
)
df_tsne <- data.frame(
  tsne1 = tsne_mvg$Y[, 1],
  tsne2 = tsne_mvg$Y[, 2],
  subtype = modify_classes(filtered_leukemia_subtypes),
  study = filtered_study_names
)
# Group rare treatment subtypes into a single 'Rare' category
df_tsne$subtype_grouped <- ifelse(
  df_tsne$subtype %in% rare_tx_subtypes,
  "Rare",
  as.character(df_tsne$subtype)
)
df_tsne$subtype_grouped <- factor(df_tsne$subtype_grouped)

p_tsne <- ggplot(df_tsne, aes(x = tsne1, y = tsne2, fill = subtype_grouped)) +
  geom_point(alpha = 1, size = 1.5, shape = 21) +
  scale_fill_brewer(palette = "Paired", name = "Subtype") +
  theme_bw() +
  labs(
    x = "t-SNE 1",
    y = "t-SNE 2",
    title = "t-SNE on MVGs (50 PCs)"
  )
p_tsne
#ggsave("../writing/figures_new/tsne_mvg_50pcs.png", p_tsne, width = 8, height = 6, dpi = 300)
```

## t-SNE on 500 subtype-informative, study-invariant genes

```{r}
# Rank genes by high eta2_subtype and low eta2_study (subtype signal, minimal batch)
df_eta2$score_subtype_vs_study <- df_eta2$eta2_subtype - df_eta2$eta2_study
idx_500 <- order(df_eta2$score_subtype_vs_study, decreasing = TRUE, na.last = TRUE)[1:500]
genes_500_subtype <- df_eta2$gene[idx_500]
col_idx_500 <- match(genes_500_subtype, counts_gene_names)

# Expression for these 500 genes only
X_500 <- scale(counts_deseq2_all[, col_idx_500])
pca_500 <- prcomp(X_500, scale. = FALSE)
pcs_50_500 <- pca_500$x[, 1:min(50, ncol(pca_500$x))]
set.seed(42)
tsne_500 <- Rtsne(
  pcs_50_500,
  dims = 2,
  perplexity = min(30, (nrow(pcs_50_500) - 1) %/% 3),
  max_iter = 1000,
  pca = FALSE
)
df_tsne_500 <- data.frame(
  tsne1 = tsne_500$Y[, 1],
  tsne2 = tsne_500$Y[, 2],
  subtype = modify_classes(filtered_leukemia_subtypes),
  study = filtered_study_names
)
# Group rare treatment subtypes into a single 'Rare' category
df_tsne_500$subtype_grouped <- ifelse(
  df_tsne_500$subtype %in% rare_tx_subtypes,
  "Rare",
  as.character(df_tsne_500$subtype)
)
df_tsne_500$subtype_grouped <- factor(df_tsne_500$subtype_grouped)

p_tsne_500 <- ggplot(df_tsne_500, aes(x = tsne1, y = tsne2, fill = subtype_grouped)) +
  geom_point(alpha = 1, size = 1.5, shape = 21) +
  scale_fill_brewer(palette = "Paired", name = "Subtype") +
  theme_bw() +
  labs(
    x = "t-SNE 1",
    y = "t-SNE 2",
    title = "t-SNE on 500 genes (high eta2 subtype, low eta2 study; 50 PCs)"
  )
p_tsne_500
#ggsave("../writing/figures_new/tsne_500_subtype_vs_study.png", p_tsne_500, width = 8, height = 6, dpi = 300)
```

# Ensemble Weights Analysis

##  --CV Weights Visualization

```{r}
# Read global and OvR ensemble weights
df_global <- read.csv("../data/out/inner_cv/ensemble_weights_merged_summed/cv/ovr_ensemble_weights_used.csv")

# Select weight columns and fold info from global weights
df_global_weights <- df_global %>%
  dplyr::select(fold, svm_weight, xgb_weight, nn_weight) %>%
  tidyr::pivot_longer(
    cols = c(svm_weight, xgb_weight, nn_weight),
    names_to = "model",
    values_to = "weight"
  )

# Combine both datasets
df_long <- df_global_weights

# Tidy model names
df_long$model <- gsub("_weight", "", df_long$model)
df_long$model <- gsub("xgb", "XGBoost", df_long$model, ignore.case = TRUE)
df_long$model <- gsub("svm", "SVM", df_long$model, ignore.case = TRUE)
df_long$model <- gsub("nn", "DNN", df_long$model, ignore.case = TRUE)

# Convert fold to factor
df_long$fold <- factor(paste0("Fold ", df_long$fold))
df_long$weight <- as.numeric(df_long$weight)

# Create boxplot with individual points
p_weights_cv <- ggplot(df_long, aes(x = model, y = weight)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(aes(color = fold), width = 0.2, size = 2, alpha = 0.6) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "CV: Ensemble weights distribution (OvR)",
    x = "Model",
    y = "Weight",
    color = "Fold"
  ) +
  ylim(0, 1)

p_weights_cv
ggsave('../writing/figures_new/weights_per_fold_cv.png', p_weights_cv, width = 10, height = 6, dpi = 300)
```

##  --LOSO Weights visualisation

```{r}
# Read global and OvR ensemble weights
df_global <- read.csv("../data/out/inner_cv/ensemble_weights_merged_summed/loso/ovr_ensemble_weights_used.csv")

# Select weight columns and fold info from global weights
df_global_weights <- df_global %>%
  dplyr::select(fold, svm_weight, xgb_weight, nn_weight) %>%
  tidyr::pivot_longer(
    cols = c(svm_weight, xgb_weight, nn_weight),
    names_to = "model",
    values_to = "weight"
  )

# Combine both datasets
df_long <- df_global_weights

# Tidy model names
df_long$model <- gsub("_weight", "", df_long$model)
df_long$model <- gsub("xgb", "XGBoost", df_long$model, ignore.case = TRUE)
df_long$model <- gsub("svm", "SVM", df_long$model, ignore.case = TRUE)
df_long$model <- gsub("nn", "DNN", df_long$model, ignore.case = TRUE)

# Convert fold to factor
df_long$fold <- factor(paste0("Fold ", df_long$fold))
df_long$weight <- as.numeric(df_long$weight)

# Create boxplot with individual points
p_weights_loso <- ggplot(df_long, aes(x = model, y = weight)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(aes(color = fold), width = 0.2, size = 2, alpha = 0.6) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "LOSO: Global ensemble weights distribution",
    x = "Model",
    y = "Weight",
    color = "Fold"
  ) +
  ylim(0, 1)

p_weights_loso
ggsave('../writing/figures_new/weights_per_fold_loso.png', p_weights_loso, width = 10, height = 6, dpi = 300)
```

# Probability Cutoffs Analysis

###  --CV Rejection Summary

```{r}
outer_merged$rejection_summary$cv$summary_stats

outer_merged$rejection_summary$loso$summary_stats
```

###  --Cutoff Visualization

```{r}
df_rej <- outer_merged$rejection_summary$cv$detailed_results

p_rej <- ggplot(df_rej) +
  geom_boxplot(aes(x = model, y = prob_cutoff)) +
  #geom_line(linewidth = 1, aes(x = model, y = weight, color = fold, group = fold)) +
  geom_point(size = 3, aes(x = model, y = prob_cutoff, color = fold, group = fold)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "CV: Probability cutoffs across models and folds",
    x = "Model",
    y = "Probability cutoff",
    color = "Fold"
  ) 
p_rej
```

# Validation Cohort Analysis (MLL Lab)

##  --Load and Process MLL Lab Data

```{r}
map_truth_to_canonical <- function(truth_labels) {
  mapping <- c(
    "AML-MR_cyto" = "AML with MDS-related cytogenetic abnormalities",
    "AML-MR_mut" = "AML with MDS-related gene mutations",
    "AML-CEBPA" = "AML with in-frame bZIP CEBPA",
    "AML-CBFB::MYH11" = "AML with inv(16)/t(16;16)/CBFB::MYH11",
    "AML-NPM1" = "AML with mutated NPM1",
    "AML-TP53" = "AML with mutated TP53",
    "AML-DEK::NUP214" = "AML with t(6;9)/DEK::NUP214",
    "AML-RUNX1::RUNX1T1" = "AML with t(8;21)/RUNX1::RUNX1T1",
    "AML-KMT2A::MLLT3" = "AML with t(9;11)/MLLT3::KMT2A",
    "APL-PML::RARA" = "APL, t(15;17)/PML::RARA",
    "AML-GATA2::MECOM" = "MECOM rearrangement",
    "AML-otherMECOM" = "MECOM rearrangement"
  )
  
  sapply(truth_labels, function(x) {
    if (x %in% names(mapping)) {
      return(mapping[x])
    } else {
      return(x)  # Return NA for unmapped labels
    }
  }, USE.NAMES = FALSE)
}

files <- list.files("../data/MLL_lab/STAR_AML_MLLlab_predictions", full.names = T)

#get_kappa <- function(file){
  MLL_lab <- read.csv("../data/MLL_lab/20231017_SampleMetadata.csv", sep = ";")
MLL_lab_truth <- MLL_lab$ICC_2022
MLL_lab_pred_df <- read.csv(files[1])
MLL_lab_truth <- map_truth_to_canonical(MLL_lab_truth)
#MLL_lab_truth[(MLL_lab$WHO_2022 == "AML-CEBPA" & MLL_lab$WHO_2017 == "AML with biallelic CEBPA" & MLL_lab$ICC_2022 != "AML-CEBPA")] <- "AML with in frame bZIP CEBPA"
MLL_lab_truth[grepl("MDS|TP53", MLL_lab_truth)] <- "AML MDS"
MLL_lab_truth[grepl("KMT2A",MLL_lab_truth) & !grepl("MLLT3", MLL_lab_truth)] <- "Other KMT2A"
# MECOM merged to same display name as elsewhere (MECOM rearrangement)
MLL_lab_truth[grepl("MECOM", MLL_lab_truth)] <- "MECOM rearrangement"
MLL_lab_truth <- make.names(MLL_lab_truth)


MLL_lab_predictions <- MLL_lab_pred_df$prediction
table(MLL_lab_predictions)
MLL_lab_predictions[grepl("MDS|TP53", MLL_lab_predictions)]<- "AML MDS"
MLL_lab_predictions[grepl("KMT2A",MLL_lab_predictions) & !grepl("MLLT3", MLL_lab_predictions)] <- "Other KMT2A"
# MECOM merged to same display name as elsewhere (MECOM rearrangement)
MLL_lab_predictions[grepl("MECOM", MLL_lab_predictions)] <- "MECOM rearrangement"
MLL_lab_predictions[grepl("KAT6A|ETV6|NUP98|CBFA2T3", MLL_lab_predictions)]<- "AML-other rare transloc"
MLL_lab_predictions <- make.names(MLL_lab_predictions)
MLL_lab_predictions <- gsub("_", ".", MLL_lab_predictions)

levels <- unique(c(MLL_lab_predictions,MLL_lab_truth))
MLL_lab_predictions <- factor(MLL_lab_predictions, levels = levels)
MLL_lab_truth <- factor(MLL_lab_truth, levels = levels)

res_MLL <- caret::confusionMatrix(MLL_lab_predictions[MLL_lab_truth != "AML..NOS"], MLL_lab_truth[MLL_lab_truth != "AML..NOS"])

round(res_MLL$overall["Kappa"], 3)
#}

#test <- data.frame(files, kappa = unlist(lapply(files, get_kappa)))
```

##  --Table 2: MLL Lab Cohort Distribution

```{r results='asis'}
# Create Table 2 for MLL lab data - matching format of Table 1
library(dplyr)
library(knitr)
library(kableExtra)

# Create data frame for MLL lab subtypes
MLL_lab_df <- data.frame(
  Subtype = MLL_lab_truth,
  Cohort = "MLL Munich",
  stringsAsFactors = FALSE
)

# Count samples per subtype
MLL_lab_counts <- as.data.frame(table(MLL_lab_df$Subtype))
colnames(MLL_lab_counts) <- c("Subtype", "MLL Munich")

# Sort by count (high to low)
MLL_lab_counts <- MLL_lab_counts[order(-MLL_lab_counts$`MLL Munich`), ]

# Clean up subtype names for display
MLL_lab_counts$Subtype <- as.character(MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("\\.", " ", MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("AML with inv 16  t 16 16  CBFB  MYH11", "CBFB::MYH11", MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("APL  t 15 17  PML  RARA", "PML::RARA", MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("AML with t 6 9  DEK  NUP214", "DEK::NUP214", MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("AML with mutated NPM1", "Mutated NPM1", MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("AML with t 8 21  RUNX1  RUNX1T1", "RUNX1::RUNX1T1", MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("AML with in frame bZIP CEBPA", "CEBPA bZIP in-frame", MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("AML with t 9 11  MLLT3  KMT2A", "MLLT3::KMT2A", MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("AML MDS or MECOM", "MDS-related and MECOM rearrangement", MLL_lab_counts$Subtype)

# Add total row
total_row <- data.frame(
  Subtype = "Total",
  `MLL Munich` = sum(MLL_lab_counts$`MLL Munich`),
  stringsAsFactors = FALSE
)
colnames(total_row) <- c("Subtype", "MLL Munich")
MLL_lab_counts_with_total <- rbind(MLL_lab_counts, total_row)

# Create formatted table matching Table 1 style
mll_table <- kable(MLL_lab_counts_with_total, 
      format = "html",
      caption = "Table 2: Distribution of Leukemia Subtypes in MLL Lab Cohort",
      row.names = FALSE, 
      escape = FALSE) %>%
  kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(nrow(MLL_lab_counts_with_total), bold = TRUE, background = "#e6f2ff") # %>%
  #add_header_above(c(" " = 1, "Adult Cohort" = 1))

mll_table
save_kable(mll_table, "../writing/tables_new/table2.png", zoom = 2)
```

##  --MLL Lab Performance Metrics

###  --Overall Performance (No Cutoff)

```{r}
res_MLL <- caret::confusionMatrix(MLL_lab_predictions[MLL_lab_truth != "AML..NOS"], MLL_lab_truth[MLL_lab_truth != "AML..NOS"])

round(res_MLL$overall, 3)
```

```{r}
res_MLL$table
```


```{r}
MLL_lab[which(MLL_lab_predictions == "AML.with.in.frame.bZIP.CEBPA" & MLL_lab_truth != "AML..NOS" & MLL_lab_truth != "AML.with.in.frame.bZIP.CEBPA"),]$WHO_2022
```


###  --Performance with Probability Cutoff 

```{r}
cutoff <- read.csv("../data/out/final_train_test/cutoffs_merged/train_test_cutoffs_merged.csv")
MLL_lab_predictions_passes <- as.logical(MLL_lab_pred_df$prediction_passed_cutoff)

res_MLL_cutoff <- caret::confusionMatrix(MLL_lab_predictions[MLL_lab_truth != "AML..NOS" & MLL_lab_predictions_passes], MLL_lab_truth[MLL_lab_truth != "AML..NOS" & MLL_lab_predictions_passes])

perc_rej <- 100 - sum(MLL_lab_predictions_passes & MLL_lab_truth != "AML..NOS") / n * 100 
round(perc_rej, 3)
round(res_MLL_cutoff$overall, 3)
```

```{r}
MLL_lab_truth[MLL_lab_truth != "AML..NOS" & !MLL_lab_predictions_passes]
```
```{r}
caret::confusionMatrix(MLL_lab_predictions[MLL_lab_truth != "AML..NOS" & !MLL_lab_predictions_passes], MLL_lab_truth[MLL_lab_truth != "AML..NOS" & !MLL_lab_predictions_passes])
```

