---
title: "Nested target-risk calibration: ensemble and rejector comparison"
output: html_document
---

Run first from repo root: `Rscript R/calibration_reject_models.R`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(readr)
  library(ggplot2)
  library(ggridges)
  library(patchwork)
  library(scales)
})

SPLIT_TYPE_LEVELS <- c("LOSO", "CV")
LABEL_DISPLAY_LEVELS <- c("Collapsed", "Collapsed (max)", "Full subtypes")
OPERATING_RISK_PCTS <- c(5, 7.5, 10)
CALIBRATION_PLOT_RISK_PCT_RANGE <- c(1, 15)
MARCE_RISK_PCT_RANGE <- c(1, 15)

filter_curve_plot_points <- function(df, loso_totals) {
  loso_ref <- loso_totals %>%
    rename(n_loso_folds_expected = n_loso_folds_total)
  df %>%
    filter(
      requested_target_risk_pct >= CALIBRATION_PLOT_RISK_PCT_RANGE[1],
      requested_target_risk_pct <= CALIBRATION_PLOT_RISK_PCT_RANGE[2]
    ) %>%
    left_join(loso_ref, by = "label_set") %>%
    filter(is.na(n_loso_folds_expected) | n_outer_folds >= n_loso_folds_expected)
}

label_set_to_display <- function(x) {
  factor(
    dplyr::recode(
      as.character(x),
      full_subtypes = "Full subtypes",
      collapsed_classes = "Collapsed",
      collapsed_maxprob = "Collapsed (max)"
    ),
    levels = LABEL_DISPLAY_LEVELS
  )
}

# Plot series scales (plot_series_levels, config_labels, config_colors) set in paths chunk.
add_plot_series <- function(df) {
  if (nrow(df) == 0L) return(df)
  if (!"class_balanced_ood" %in% names(df)) {
    return(df %>%
      mutate(
        plot_series_key = as.character(config_key),
        plot_series_label = as.character(config_label)
      ))
  }
  df %>%
    mutate(
      plot_series_key = case_when(
        is.na(class_balanced_ood) ~ as.character(config_key),
        class_balanced_ood ~ paste(as.character(config_key), "balanced", sep = "|"),
        TRUE ~ paste(as.character(config_key), "unweighted", sep = "|")
      ),
      plot_series_label = case_when(
        is.na(class_balanced_ood) ~ as.character(config_label),
        class_balanced_ood ~ paste0(as.character(config_label), " | OOD wt: balanced"),
        TRUE ~ paste0(as.character(config_label), " | OOD wt: unweighted")
      )
    )
}

build_plot_series_meta <- function(df) {
  meta_cols <- c("plot_series_key", "plot_series_label", "config_key")
  if ("class_balanced_ood" %in% names(df)) {
    meta_cols <- c(meta_cols, "class_balanced_ood")
  }
  df %>%
    add_plot_series() %>%
    distinct(across(all_of(meta_cols))) %>%
    mutate(
      is_maxprob = grepl("^maxprob", config_key),
      ood_sort = if ("class_balanced_ood" %in% names(.)) {
        dplyr::case_when(
          is.na(class_balanced_ood) ~ 0L,
          class_balanced_ood ~ 1L,
          TRUE ~ 2L
        )
      } else {
        0L
      }
    ) %>%
    arrange(is_maxprob, config_key, ood_sort)
}

init_plot_series_scales <- function(series_meta) {
  plot_series_levels <<- series_meta$plot_series_key
  config_labels <<- setNames(series_meta$plot_series_label, series_meta$plot_series_key)
  config_colors <<- setNames(
    scales::hue_pal(l = 65, c = 100)(nrow(series_meta)),
    series_meta$plot_series_key
  )
  invisible(series_meta)
}

THRESHOLD_METHOD_LEVELS <- c("pooled_oof", "jackknife_adjusted", "ucb_95")
THRESHOLD_METHOD_LABELS <- c(
  "Pooled out-of-fold",
  "Jackknife-adjusted",
  "UCB (95%, one-sided)"
)
names(THRESHOLD_METHOD_LABELS) <- THRESHOLD_METHOD_LEVELS
threshold_method_linetypes <- c("solid", "dashed", "dotdash")
names(threshold_method_linetypes) <- THRESHOLD_METHOD_LEVELS

# Final deployment rejectors (SVM-only; single-head max-prob and ridge in-model).
DEPLOY_REJECTOR_LEVELS <- c(
  "svm_single_head",
  "svm_ridge_in_model"
)
DEPLOY_REJECTOR_LABELS <- c(
  "SVM max-prob | Single-head",
  "SVM ridge (in-model) | Single-head"
)
names(DEPLOY_REJECTOR_LABELS) <- DEPLOY_REJECTOR_LEVELS
deploy_rejector_colors <- c("#ff7f0e", "#2ca02c")
names(deploy_rejector_colors) <- DEPLOY_REJECTOR_LEVELS
```

SVM-only calibration compares two single-head rejectors: **max-prob** and **ridge (in-model confidence features)**. Threshold methods: **pooled out-of-fold**, **jackknife-adjusted**, and **UCB 95% (one-sided)**.

**Threshold calibration (three methods, reported side by side):**

- **Pooled out-of-fold** (default): pick the probability cutoff from pooled out-of-fold scores on the outer train pool.
- **Jackknife-adjusted**: within the pool, repeatedly hold out one study, calibrate on the rest, measure realized minus requested risk on the held-out study; average that gap and subtract it from the requested risk before choosing the cutoff on the full pool.
- **UCB (95%, one-sided)**: pick the most permissive cutoff whose one-sided 95% Wilson upper bound on accepted error is <= requested risk.

Run from repo root: `Rscript R/calibration_reject_models.R`

```{r paths}
repo_root <- if (file.exists("R/outer_cv_analysis.R")) "." else ".."
out_dir <- file.path(repo_root, "data/out/outer_cv/calibration_feature_utility_selection_safe")

summary_operating_path <- file.path(out_dir, "nested_target_risk_summary_operating.csv")
marce_path <- file.path(out_dir, "nested_target_risk_marce.csv")
calibration_curve_path <- file.path(out_dir, "nested_target_risk_calibration_curve.csv")
calibration_fold_coverage_path <- file.path(out_dir, "nested_target_risk_calibration_fold_coverage.csv")
per_fold_operating_path <- file.path(out_dir, "nested_target_risk_per_fold_operating.csv")
probability_samples_path <- file.path(out_dir, "nested_target_risk_probability_samples.csv")
inner_logloss_summary_path <- file.path(out_dir, "nested_target_risk_inner_logloss_summary.csv")

for (p in c(
  summary_operating_path, marce_path, calibration_curve_path, probability_samples_path
)) {
  if (!file.exists(p)) {
    stop("Missing ", p, ". Run: Rscript R/calibration_reject_models.R")
  }
}

summary_operating_df <- read_csv(summary_operating_path, show_col_types = FALSE) %>%
  filter(requested_target_risk_pct %in% OPERATING_RISK_PCTS) %>%
  add_plot_series()

init_plot_series_scales(build_plot_series_meta(summary_operating_df))

marce_df <- read_csv(marce_path, show_col_types = FALSE) %>% add_plot_series()
calibration_curve_df <- read_csv(calibration_curve_path, show_col_types = FALSE) %>% add_plot_series()
per_fold_operating_df <- read_csv(per_fold_operating_path, show_col_types = FALSE) %>% add_plot_series()
fold_coverage_df <- if (file.exists(calibration_fold_coverage_path)) {
  read_csv(calibration_fold_coverage_path, show_col_types = FALSE) %>% add_plot_series()
} else {
  data.frame()
}
probability_samples_df <- read_csv(probability_samples_path, show_col_types = FALSE) %>% add_plot_series()
inner_logloss_summary_df <- if (file.exists(inner_logloss_summary_path)) {
  read_csv(inner_logloss_summary_path, show_col_types = FALSE)
} else {
  data.frame()
}
```

## Inner CV log loss

Mean inner cross-validation log loss per outer fold (study-blocked inner CV on the outer train pool), averaged across LOSO folds. Lower is better. **Max-prob** uses a fixed `max_prob` GLM (no feature selection); **ridge** uses the inner winner (`inner_rank == 1`). Rank is within `label_set × split_type × nested_rejector_mode`.

```{r inner_logloss_table}
if (nrow(inner_logloss_summary_df) == 0L) {
  knitr::kable(
    data.frame(note = "Run R/calibration_reject_models.R to generate nested_target_risk_inner_logloss_summary.csv"),
    caption = "Inner CV log loss (not yet available)"
  )
} else {
  inner_logloss_display <- inner_logloss_summary_df %>%
    mutate(
      split_type = toupper(as.character(split_type)),
      label_display = label_set_to_display(label_set),
      nested_rejector_mode = dplyr::recode(
        as.character(nested_rejector_mode),
        single_head = "Single-head",
        two_head_min = "Two-head (min nested)"
      )
    ) %>%
    select(
      rank_mean_logloss, config_label, ensemble_key, recipe_key, rejector_method,
      nested_rejector_mode, label_display, split_type,
      n_outer_folds, mean_inner_logloss, sd_inner_logloss, inner_winner_alpha
    ) %>%
    arrange(label_display, split_type, nested_rejector_mode, rank_mean_logloss)

  knitr::kable(
    inner_logloss_display,
    digits = 4,
    col.names = c(
      "Rank", "Config", "Ensemble", "Recipe", "Rejector", "Head mode",
      "Label set", "Split", "N folds", "Mean log loss", "SD log loss", "Winner alpha"
    ),
    caption = "Inner CV log loss summary (max-prob and ridge jobs)"
  )
}
```

## Operating points (5%, 7.5%, 10% requested risk)

Mean across outer folds when **all** LOSO folds have a successful eval at that risk; otherwise means and SDs are **NA**. Lower risk gap and higher coverage are better. Both threshold methods are shown.

```{r operating_table}
loso_fold_totals_op <- per_fold_operating_df %>%
  filter(toupper(as.character(split_type)) == "LOSO") %>%
  group_by(label_set) %>%
  summarise(n_loso_folds_total_ref = dplyr::n_distinct(target_fold), .groups = "drop")

operating_display <- summary_operating_df %>%
  left_join(loso_fold_totals_op, by = "label_set") %>%
  mutate(
    split_type = toupper(as.character(split_type)),
    label_display = label_set_to_display(label_set),
    plot_series_key = factor(as.character(plot_series_key), levels = plot_series_levels),
    threshold_method = factor(
      if ("threshold_method" %in% names(.)) as.character(threshold_method) else "pooled_oof",
      levels = THRESHOLD_METHOD_LEVELS,
      labels = THRESHOLD_METHOD_LABELS
    ),
    mean_abs_risk_gap_pct = 100 * mean_abs_risk_gap,
    mean_outer_coverage_seen_pct = 100 * mean_outer_coverage_seen,
    mean_outer_risk_pct = 100 * mean_outer_risk_all_accepted,
    sd_outer_risk_pct = 100 * sd_outer_risk_all_accepted,
    mean_jackknife_gap_pct = if ("mean_jackknife_gap" %in% names(.)) 100 * mean_jackknife_gap else NA_real_,
    n_loso_folds_total = dplyr::coalesce(n_loso_folds_total, n_loso_folds_total_ref),
    all_outer_folds = if ("all_outer_folds" %in% names(.)) {
      all_outer_folds
    } else {
      !is.na(n_loso_folds_total) & (n_outer_folds == n_loso_folds_total)
    }
  ) %>%
  select(
    plot_series_label, label_display, split_type, threshold_method, requested_target_risk_pct,
    n_outer_folds, n_loso_folds_total, all_outer_folds,
    mean_abs_risk_gap_pct, mean_outer_risk_pct, sd_outer_risk_pct,
    mean_outer_coverage_seen_pct, mean_jackknife_gap_pct
  ) %>%
  arrange(plot_series_label, label_display, split_type, threshold_method, requested_target_risk_pct)

knitr::kable(operating_display, digits = 2, caption = "Outer-fold means at 5%, 7.5%, and 10% (NA unless all folds succeed)")
```

## Outer-fold coverage for calibration curves

RR/RC curves plot **1–15%** (0.1% steps). A point is included only when **all** LOSO outer folds have a successful eval at that requested risk.

```{r curve_fold_summary}
loso_fold_totals <- per_fold_operating_df %>%
  filter(toupper(as.character(split_type)) == "LOSO") %>%
  group_by(label_set) %>%
  summarise(n_loso_folds_total = dplyr::n_distinct(target_fold), .groups = "drop")

curve_series_summary <- calibration_curve_df %>%
  filter(
    requested_target_risk_pct >= CALIBRATION_PLOT_RISK_PCT_RANGE[1],
    requested_target_risk_pct <= CALIBRATION_PLOT_RISK_PCT_RANGE[2]
  ) %>%
  mutate(
    label_display = label_set_to_display(label_set),
    plot_series_key = factor(as.character(plot_series_key), levels = plot_series_levels),
    config_label = if ("plot_series_label" %in% names(.)) {
      as.character(plot_series_label)
    } else if ("config_label" %in% names(.)) {
      as.character(config_label)
    } else {
      unname(config_labels[as.character(plot_series_key)])
    },
    threshold_method = factor(as.character(threshold_method), levels = THRESHOLD_METHOD_LEVELS, labels = THRESHOLD_METHOD_LABELS)
  ) %>%
  group_by(config_label, label_display, label_set, split_type, threshold_method) %>%
  summarise(
    n_risk_grid_points_plotted = dplyr::n(),
    n_outer_folds_min = min(n_outer_folds, na.rm = TRUE),
    n_outer_folds_max = max(n_outer_folds, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(loso_fold_totals, by = "label_set") %>%
  mutate(
    label_display = factor(label_display, levels = LABEL_DISPLAY_LEVELS),
    all_folds_at_every_plotted_point = n_outer_folds_min == n_loso_folds_total &
      n_outer_folds_max == n_loso_folds_total
  ) %>%
  arrange(config_label, label_display, threshold_method)

knitr::kable(
  curve_series_summary %>%
    select(
      config_label, label_display, threshold_method,
      n_risk_grid_points_plotted, n_outer_folds_min, n_outer_folds_max,
      n_loso_folds_total, all_folds_at_every_plotted_point
    ) %>%
    mutate(label_display = as.character(label_display)),
  digits = 1,
  caption = "Plotted RR/RC points (1–15%, 0.1% steps); each point requires all LOSO folds"
)
```

```{r fold_coverage_diagnostic, fig.width=11, fig.height=8}
if (nrow(fold_coverage_df) > 0L) {
  fold_cov_plot_df <- fold_coverage_df %>% filter( !grepl("unweighted", plot_series_key)) %>%
    mutate(
      label_display = label_set_to_display(label_set),
      plot_series_key = factor(as.character(plot_series_key), levels = plot_series_levels),
      threshold_method = factor(as.character(threshold_method), levels = THRESHOLD_METHOD_LEVELS, labels = THRESHOLD_METHOD_LABELS)
    )
  p_fold_cov <- ggplot(
    fold_cov_plot_df,
    aes(
      x = requested_target_risk_pct,
      y = n_outer_folds_available,
      color = plot_series_key,
      group = plot_series_key
    )
  ) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 1.1) +
    geom_hline(
      aes(yintercept = n_loso_folds_total),
      linetype = "dotted",
      color = "grey40",
      linewidth = 0.5
    ) +
    scale_color_manual(values = config_colors, labels = config_labels, name = NULL) +
    facet_grid(label_display ~ threshold_method) +
    scale_x_continuous(breaks = seq(2, 15, by = 2)) +
    labs(
      title = "Outer folds with successful eval per requested risk (1–15% grid, 0.1% steps)",
      subtitle = "Dotted line = total LOSO folds; points below the line are omitted from RR/RC curves",
      x = "Requested risk (%)",
      y = "Outer folds available"
    ) +
    theme_bw() +
    theme(legend.position = "bottom", legend.text = element_text(size = 8))
  print(p_fold_cov)
} else {
  operating_fold_coverage <- per_fold_operating_df %>%
    filter(toupper(as.character(split_type)) == "LOSO") %>%
    mutate(
      label_display = label_set_to_display(label_set),
      plot_series_key = factor(as.character(plot_series_key), levels = plot_series_levels),
      threshold_method = factor(as.character(threshold_method), levels = THRESHOLD_METHOD_LEVELS, labels = THRESHOLD_METHOD_LABELS)
    ) %>%
    group_by(plot_series_key, label_display, threshold_method, requested_target_risk_pct) %>%
    summarise(n_outer_folds_available = dplyr::n(), .groups = "drop") %>%
    group_by(plot_series_key, label_display, threshold_method) %>%
    mutate(n_loso_folds_total = max(n_outer_folds_available)) %>%
    ungroup()
  p_fold_cov_op <- ggplot(
    operating_fold_coverage,
    aes(x = requested_target_risk_pct, y = n_outer_folds_available, color = plot_series_key, group = plot_series_key)
  ) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_hline(aes(yintercept = n_loso_folds_total), linetype = "dotted", color = "grey40") +
    #scale_color_manual(values = config_colors, labels = config_labels, name = NULL) +
    facet_grid(label_display ~ threshold_method) +
    labs(
      title = "Outer folds with successful eval (operating risks: 5%, 7.5%, 10%)",
      subtitle = "Re-run calibration_reject_models.R for full 1–15% fold-coverage export",
      x = "Requested risk (%)",
      y = "Outer folds available"
    ) +
    theme_bw() +
    theme(legend.position = "bottom")+ scale_color_brewer(palette = "Set3")
  print(p_fold_cov_op)
}
```

## Mean Absolute Risk Calibration Error (MARCE)

MARCE = mean of **|realized risk − requested risk|** over **1–15%** (0.1% steps), averaging only grid points where all LOSO outer folds evaluated successfully.

```{r marce_table}
marce_display <- marce_df %>%
  mutate(
    split_type = toupper(as.character(split_type)),
    label_display = label_set_to_display(label_set),
    plot_series_key = factor(as.character(plot_series_key), levels = plot_series_levels),
    plot_series_label = if ("plot_series_label" %in% names(.)) {
      as.character(plot_series_label)
    } else if ("config_label" %in% names(.)) {
      as.character(config_label)
    } else {
      unname(config_labels[as.character(plot_series_key)])
    },
    threshold_method = factor(
      if ("threshold_method" %in% names(.)) as.character(threshold_method) else "pooled_oof",
      levels = THRESHOLD_METHOD_LEVELS,
      labels = THRESHOLD_METHOD_LABELS
    ),
    ood_weight_variant = if ("class_balanced_ood" %in% names(.)) {
      dplyr::case_when(
        is.na(class_balanced_ood) ~ NA_character_,
        class_balanced_ood ~ "balanced",
        TRUE ~ "unweighted"
      )
    } else {
      NA_character_
    }
  ) %>%
  select(
    plot_series_label, ood_weight_variant, label_display, split_type, threshold_method,
    n_risk_grid_points, n_outer_folds_curve, marce_pct
  ) %>%
  arrange(plot_series_label, label_display, split_type, threshold_method)

knitr::kable(marce_display, digits = 3)
```

```{r marce_bars, fig.width=12, fig.height=6}
marce_bar_df <- marce_df %>% filter(!grepl("unweighted", plot_series_key)) %>%
  left_join(loso_fold_totals_op, by = "label_set") %>%
  mutate(
    split_type_fac = factor(toupper(as.character(split_type)), levels = SPLIT_TYPE_LEVELS),
    label_display = label_set_to_display(label_set),
    plot_series_key = factor(as.character(plot_series_key), levels = plot_series_levels),
    plot_series_label = if ("plot_series_label" %in% names(.)) {
      as.character(plot_series_label)
    } else if ("config_label" %in% names(.)) {
      as.character(config_label)
    } else {
      unname(config_labels[as.character(plot_series_key)])
    },
    threshold_method = factor(
      if ("threshold_method" %in% names(.)) as.character(threshold_method) else "pooled_oof",
      levels = THRESHOLD_METHOD_LEVELS,
      labels = THRESHOLD_METHOD_LABELS
    ),
    n_outer_folds_marce = dplyr::coalesce(
      if ("n_outer_folds_marce" %in% names(.)) .data$n_outer_folds_marce else NA_real_,
      if ("n_outer_folds_curve" %in% names(.)) .data$n_outer_folds_curve else NA_real_
    ),
    n_loso_folds_total = dplyr::coalesce(
      if ("n_loso_folds_total" %in% names(.)) .data$n_loso_folds_total else NA_real_,
      .data$n_loso_folds_total_ref
    )
  ) 

p_marce <- ggplot(
  marce_bar_df %>%
    filter(
      split_type_fac == "LOSO",
      is.na(n_loso_folds_total) | n_outer_folds_marce >= n_loso_folds_total
    ),
  aes(x = plot_series_key, y = marce_pct, fill = threshold_method)
) +
  geom_col(position = position_dodge(width = 0.75), width = 0.65) +
  scale_x_discrete(labels = config_labels) +
 # scale_fill_manual(values = c("#bdbdbd", "#636363"), name = "Threshold") +
  facet_wrap(~ label_display, ncol = 2) +
  labs(
    title = "MARCE on 1–15% requested-risk grid (LOSO)",
    subtitle = "Mean |realized − requested| risk (%); per-point inclusion requires all LOSO folds",
    x = NULL,
    y = "MARCE (%)"
  ) +
  theme_bw() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 25, hjust = 1))

print(p_marce)
```

## Requested vs realized risk (1–15% grid, 0.1% steps, all folds required per point)

```{r calibration_curves_risk, fig.width=11, fig.height=10}
curve_plot_df <- filter_curve_plot_points(calibration_curve_df, loso_fold_totals) %>%
  mutate(
    split_type_fac = factor(toupper(split_type), levels = SPLIT_TYPE_LEVELS),
    label_display = label_set_to_display(label_set),
    plot_series_key = factor(as.character(plot_series_key), levels = plot_series_levels),
    threshold_method = factor(
      if ("threshold_method" %in% names(.)) as.character(threshold_method) else "pooled_oof",
      levels = THRESHOLD_METHOD_LEVELS
    )
  ) %>% #filter(threshold_method != "jackknife_adjusted" & !grepl("unweighted", plot_series_key)) %>%
  filter(grepl("svm", config_key)) %>%
  arrange(label_display, split_type_fac, plot_series_key, threshold_method, requested_target_risk_pct)

curve_fold_labelss <- curve_plot_df %>%
  group_by(label_display, split_type_fac, plot_series_key, threshold_method) %>%
  slice_max(requested_target_risk_pct, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(label_n = paste0("n=", n_outer_folds))

p_rr <- ggplot(
  curve_plot_df %>% filter(grepl("svm_ridge_in_model", config_key)),
  aes(
    x = requested_target_risk_pct,
    y = realized_risk_pct,
    color = threshold_method,
    linetype = threshold_method,
    group = interaction(plot_series_key, threshold_method)
  )
) +
  geom_line(linewidth = 0.9) +
  #geom_point(size = 1.2) +
  geom_text(
    data = curve_fold_labels,
    aes(label = label_n),
    hjust = -0.08,
    vjust = -0.3,
    size = 2.2,
    show.legend = FALSE
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey45") +
  #scale_color_manual(values = config_colors, labels = config_labels, name = NULL) +
  scale_linetype_manual(values = threshold_method_linetypes, labels = THRESHOLD_METHOD_LABELS, name = "Threshold") +
  facet_grid(label_display ~ split_type_fac) +
  coord_cartesian(xlim = c(0, 15.5), ylim = c(0, 20)) +
  labs(
    title = "Requested vs realized accepted risk",
    subtitle = "1–15% grid (0.1% steps); points omitted unless all LOSO folds eval successfully (n at right)",
    x = "Requested risk (%)",
    y = "Realized accepted risk (%)"
  ) +
  theme_bw() +
  theme(legend.position = "bottom", legend.text = element_text(size = 6)) + scale_color_brewer(palette = "Set3")

print(p_rr)
```

## Requested risk vs realized seen-class coverage

```{r calibration_curves_coverage, fig.width=11, fig.height=10}
p_rc <- ggplot(
  curve_plot_df,
  aes(
    x = realized_risk_pct,
    y = realized_coverage_seen_pct,
    color = plot_series_key,
    #linetype = threshold_method,
    group = interaction(plot_series_key)#, threshold_method)
  )
) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.2) +
  geom_text(
    data = curve_fold_labels,
    aes(label = label_n),
    hjust = -0.08,
    vjust = -0.3,
    size = 2.2,
    show.legend = FALSE
  ) +
  #scale_color_manual(values = config_colors, labels = config_labels, name = NULL) +
  #scale_linetype_manual(values = threshold_method_linetypes, labels = THRESHOLD_METHOD_LABELS, name = "Threshold") +
  facet_grid(label_display ~ split_type_fac) +
  coord_cartesian(xlim = c(0, 15.5)) +
  labs(
    title = "Realized risk vs realized seen-class coverage",
    subtitle = "1–15% grid (0.1% steps); points omitted unless all LOSO folds eval successfully (n at right)",
    x = "Realized risk (%)",
    y = "Realized seen-class coverage (%)"
  ) +
  theme_bw() +
  theme(legend.position = "bottom", legend.text = element_text(size = 6)) + scale_color_brewer(palette = "Set3")

print(p_rc)
```

## Risk gap and coverage at operating points (5%, 7.5%, 10%)

```{r operating_bars, fig.width=12, fig.height=8}
op_bar_df <- summary_operating_df %>%
  left_join(loso_fold_totals_op, by = "label_set") %>%
  mutate(
    split_type_fac = factor(toupper(split_type), levels = SPLIT_TYPE_LEVELS),
    label_display = label_set_to_display(label_set),
    plot_series_key = factor(as.character(plot_series_key), levels = plot_series_levels),
    threshold_method = factor(
      if ("threshold_method" %in% names(.)) as.character(threshold_method) else "pooled_oof",
      levels = THRESHOLD_METHOD_LEVELS,
      labels = THRESHOLD_METHOD_LABELS
    ),
    n_loso_folds_total = dplyr::coalesce(n_loso_folds_total, n_loso_folds_total_ref),
    all_outer_folds = if ("all_outer_folds" %in% names(.)) {
      all_outer_folds
    } else {
      !is.na(n_loso_folds_total) & (n_outer_folds == n_loso_folds_total)
    },
    operating_risk_pct = factor(requested_target_risk_pct, levels = OPERATING_RISK_PCTS)
  )

op_bar_plot_df <- op_bar_df %>%
  filter(split_type_fac == "LOSO", all_outer_folds)

p_gap <- ggplot(
  op_bar_plot_df,
  aes(x = plot_series_key, y = 100 * mean_outer_risk_all_accepted, fill = threshold_method)
) +
  geom_col(position = position_dodge(width = 0.75), width = 0.65) +
  scale_x_discrete(labels = config_labels) +
  scale_fill_manual(values = c("#bdbdbd", "#636363"), name = "Threshold") +
  facet_grid(operating_risk_pct ~ label_display, labeller = labeller(operating_risk_pct = function(x) paste0(x, "%"))) +
  labs(
    title = "Mean realized risk (LOSO)",
    subtitle = "Bars shown only when all LOSO folds succeed at that requested risk",
    x = NULL,
    y = "Mean absolute risk gap (%)"
  ) +
  theme_bw() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 25, hjust = 1))

p_cov <- ggplot(
  op_bar_plot_df,
  aes(x = plot_series_key, y = 100 * mean_outer_coverage_seen, fill = threshold_method)
) +
  geom_col(position = position_dodge(width = 0.75), width = 0.65) +
  scale_x_discrete(labels = config_labels) +
  scale_fill_manual(values = c("#bdbdbd", "#636363"), name = "Threshold") +
  facet_grid(operating_risk_pct ~ label_display, labeller = labeller(operating_risk_pct = function(x) paste0(x, "%"))) +
  labs(
    title = "Realized seen-class coverage at operating targets (LOSO)",
    subtitle = "Bars shown only when all LOSO folds succeed at that requested risk",
    x = NULL,
    y = "Mean coverage (%)"
  ) +
  theme_bw() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 25, hjust = 1))

print(p_gap + p_cov)
```

## Accept probability by outcome (correct, incorrect, OOD)

Densities of max-prob rejector accept probability on **LOSO** held-out folds. Ideal separation: **Correct** high, **Incorrect** and **OOD** low.

```{r prob_distributions, fig.width=12, fig.height=10}
outcome_colors <- c(
  "Incorrect" = "#d95f02",
  "OOD" = "#7570b3",
  "Correct" = "#1b9e77"
)

prob_plot_df <- probability_samples_df %>% filter(grepl("svm", config_key)) %>%
  mutate(
    outcome = case_when(
      is_seen == 0L ~ "OOD",
      correct == 1L ~ "Correct",
      correct == 0L ~ "Incorrect",
      TRUE ~ NA_character_
    ),
    outcome = factor(outcome, levels = c("Incorrect", "OOD", "Correct")),
    label_display = label_set_to_display(label_set),
    plot_series_key = factor(as.character(plot_series_key), levels = plot_series_levels)
  ) %>%
  filter(!is.na(outcome), toupper(split_type) == "LOSO")

p_probs_faceted <- ggplot(
  prob_plot_df,
  aes(x = p_hat, y = outcome, fill = outcome, color = outcome)
) +
  geom_density_ridges(
    scale = 1.8,
    alpha = 0.45,
    rel_min_height = -0.01,
    bounds = c(0, 1),
    linewidth = 0.3
  ) +
  scale_x_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = outcome_colors, guide = "none") +
  scale_color_manual(values = outcome_colors, guide = "none") +
  facet_grid(label_display ~ plot_series_key, labeller = labeller(
    plot_series_key = config_labels
  )) +
  labs(
    title = "Accept probability by outcome (all rejector configurations)",
    x = "Accept probability (p_hat)",
    y = NULL
  ) +
  theme_ridges() +
  theme(
    strip.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 9)
  )

print(p_probs_faceted)
```

### Side-by-side boxplots (LOSO, collapsed classes)

```{r prob_boxplots, fig.width=11, fig.height=5}
p_box <- ggplot(
  prob_plot_df %>% filter(label_display == "Collapsed"),
  aes(x = plot_series_key, y = p_hat, fill = outcome)
) +
  geom_boxplot(outlier.alpha = 0.25, position = position_dodge(width = 0.78), width = 0.7) +
  scale_x_discrete(labels = config_labels) +
  scale_fill_manual(values = outcome_colors, name = "Outcome") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    title = "Accept probability distributions (LOSO, collapsed classes)",
    x = NULL,
    y = "Accept probability (p_hat)"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position = "bottom")

print(p_box)
```

---

# Final deployment calibration curves

Run after `R/train_test_analysis.R`: `Rscript R/calibration_reject_models_final.R`

These curves use the **same stack as `python/predict_new_samples.py`**: pooled max-prob GLM rejectors on all training data, jackknife-adjusted thresholds, and LOSO pseudo-outer evaluation (each study held out once). Use these for the **main deployment figure** and MLL-lab interpretation. Keep the outer-CV curves above for **method development** (nested selection, rejector comparison, uncertainty bands across the full nested pipeline).

```{r final_deploy_paths}
final_out_dir <- file.path(repo_root, "data/out/final_train_test/calibration_feature_utility_selection_safe")
final_calibration_curve_path <- file.path(final_out_dir, "final_deploy_calibration_curve.csv")
final_calibration_per_fold_path <- file.path(final_out_dir, "final_deploy_calibration_curve_per_fold.csv")

if (!file.exists(final_calibration_curve_path)) {
  manifest_path <- file.path(final_out_dir, "final_deployment_manifest.csv")
  has_new_manifest <- file.exists(manifest_path) &&
    "deploy_risk_grid" %in% readr::read_csv(manifest_path, show_col_types = FALSE)$key
  deploy_curve_glob <- list.files(
    file.path(repo_root, "data/out/final_train_test"),
    pattern = "^deploy_risk_coverage_curve_maxprob_.*\\.csv$",
    recursive = TRUE,
    full.names = TRUE
  )
  hint <- if (length(deploy_curve_glob) == 0L && !has_new_manifest) {
    paste0(
      "Your manifest/cutoffs look like an older run (cutoffs only, no deploy risk-coverage curves). ",
      "Re-run from repo root with the current script:\n",
      "  Rscript R/calibration_reject_models_final.R\n",
      "Expected new outputs: final_deploy_calibration_curve.csv and ",
      "cutoffs_*/deploy_risk_coverage_curve_maxprob_*.csv"
    )
  } else {
    paste0(
      "Re-run from repo root:\n",
      "  Rscript R/calibration_reject_models_final.R"
    )
  }
  stop("Missing ", final_calibration_curve_path, ".\n", hint)
}

final_calibration_curve_df <- read_csv(final_calibration_curve_path, show_col_types = FALSE) %>%
  filter(as.character(rejector_key) %in% DEPLOY_REJECTOR_LEVELS)

final_calibration_per_fold_df <- if (file.exists(final_calibration_per_fold_path)) {
  read_csv(final_calibration_per_fold_path, show_col_types = FALSE) %>%
    filter(as.character(rejector_key) %in% DEPLOY_REJECTOR_LEVELS)
} else {
  data.frame()
}

final_loso_fold_totals <- final_calibration_curve_df %>%
  group_by(label_set) %>%
  summarise(n_loso_folds_total = max(n_outer_folds, na.rm = TRUE), .groups = "drop")

filter_final_curve_plot_points <- function(df, loso_totals) {
  loso_ref <- loso_totals %>%
    rename(n_loso_folds_expected = n_loso_folds_total)
  df %>%
    filter(
      requested_target_risk_pct >= CALIBRATION_PLOT_RISK_PCT_RANGE[1],
      requested_target_risk_pct <= CALIBRATION_PLOT_RISK_PCT_RANGE[2]
    ) %>%
    left_join(loso_ref, by = "label_set") %>%
    filter(is.na(n_loso_folds_expected) | n_outer_folds >= n_loso_folds_expected)
}
```

```{r final_deploy_curve_summary}
final_curve_series_summary <- final_calibration_curve_df %>%
  filter(
    requested_target_risk_pct >= CALIBRATION_PLOT_RISK_PCT_RANGE[1],
    requested_target_risk_pct <= CALIBRATION_PLOT_RISK_PCT_RANGE[2]
  ) %>%
  mutate(
    label_display = label_set_to_display(label_set),
    rejector_key = factor(as.character(rejector_key), levels = DEPLOY_REJECTOR_LEVELS),
    rejector_label = unname(DEPLOY_REJECTOR_LABELS[as.character(rejector_key)])
  ) %>%
  group_by(rejector_label, label_display, label_set) %>%
  summarise(
    n_risk_grid_points_plotted = dplyr::n(),
    n_outer_folds_min = min(n_outer_folds, na.rm = TRUE),
    n_outer_folds_max = max(n_outer_folds, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(final_loso_fold_totals, by = "label_set") %>%
  mutate(
    label_display = factor(label_display, levels = LABEL_DISPLAY_LEVELS),
    all_folds_at_every_plotted_point = n_outer_folds_min == n_loso_folds_total &
      n_outer_folds_max == n_loso_folds_total
  ) %>%
  arrange(rejector_label, label_display)

knitr::kable(
  final_curve_series_summary %>%
    select(
      rejector_label, label_display,
      n_risk_grid_points_plotted, n_outer_folds_min, n_outer_folds_max,
      n_loso_folds_total, all_folds_at_every_plotted_point
    ) %>%
    mutate(label_display = as.character(label_display)),
  digits = 1,
  caption = "Final deployment RR/RC points (1–15%, 0.1% steps); jackknife-adjusted thresholds"
)
```

## Final deployment: requested vs realized risk

```{r final_deploy_rr, fig.width=9, fig.height=6}
final_curve_plot_df <- filter_final_curve_plot_points(
  final_calibration_curve_df, final_loso_fold_totals
) %>%
  mutate(
    label_display = label_set_to_display(label_set),
    rejector_key = factor(as.character(rejector_key), levels = DEPLOY_REJECTOR_LEVELS)
  ) %>%
  arrange(label_display, rejector_key, requested_target_risk_pct)

final_curve_fold_labels <- final_curve_plot_df %>%
  group_by(label_display, rejector_key) %>%
  slice_max(requested_target_risk_pct, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(label_n = paste0("n=", n_outer_folds))

p_final_rr <- ggplot(
  final_curve_plot_df,
  aes(
    x = requested_target_risk_pct,
    y = realized_risk_pct,
    color = rejector_key,
    group = rejector_key
  )
) +
  geom_ribbon(
    aes(ymin = realized_risk_ci95_lo_pct, ymax = realized_risk_ci95_hi_pct, fill = rejector_key),
    alpha = 0.15,
    color = NA
  ) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.2) +
  geom_text(
    data = final_curve_fold_labels,
    aes(label = label_n),
    hjust = -0.08,
    vjust = -0.3,
    size = 2.4,
    show.legend = FALSE
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey45") +
  scale_color_manual(values = deploy_rejector_colors, labels = DEPLOY_REJECTOR_LABELS, name = NULL) +
  scale_fill_manual(values = deploy_rejector_colors, guide = "none") +
  facet_wrap(~ label_display, ncol = 2) +
  coord_cartesian(xlim = c(0, 15.5), ylim = c(0, 20)) +
  labs(
    title = "Final deployment: requested vs realized accepted risk",
    subtitle = "Pooled max-prob GLM + jackknife thresholds; LOSO pseudo-outer eval (n at right)",
    x = "Requested risk (%)",
    y = "Realized accepted risk (%)"
  ) +
  theme_bw() +
  theme(legend.position = "bottom", legend.text = element_text(size = 8))

print(p_final_rr)
```

## Final deployment: requested risk vs seen-class coverage

```{r final_deploy_rc, fig.width=9, fig.height=6}
p_final_rc <- ggplot(
  final_curve_plot_df,
  aes(
    x = requested_target_risk_pct,
    y = realized_coverage_seen_pct,
    color = rejector_key,
    group = rejector_key
  )
) +
  geom_ribbon(
    aes(
      ymin = realized_coverage_seen_ci95_lo_pct,
      ymax = realized_coverage_seen_ci95_hi_pct,
      fill = rejector_key
    ),
    alpha = 0.15,
    color = NA
  ) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.2) +
  geom_text(
    data = final_curve_fold_labels,
    aes(label = label_n),
    hjust = -0.08,
    vjust = -0.3,
    size = 2.4,
    show.legend = FALSE
  ) +
  scale_color_manual(values = deploy_rejector_colors, labels = DEPLOY_REJECTOR_LABELS, name = NULL) +
  scale_fill_manual(values = deploy_rejector_colors, guide = "none") +
  facet_wrap(~ label_display, ncol = 2) +
  coord_cartesian(xlim = c(0, 15.5)) +
  labs(
    title = "Final deployment: requested risk vs realized seen-class coverage",
    subtitle = "Pooled max-prob GLM + jackknife thresholds; LOSO pseudo-outer eval (n at right)",
    x = "Requested risk (%)",
    y = "Realized seen-class coverage (%)"
  ) +
  theme_bw() +
  theme(legend.position = "bottom", legend.text = element_text(size = 8))

print(p_final_rc)
```

```{r save_figures, include=FALSE}
base <- file.path(repo_root, "writing/figures_new/figure_rej")
dir.create(base, recursive = TRUE, showWarnings = FALSE)
ggsave(file.path(base, "rr_curve_rejector_compare.svg"), p_rr, height = 5, width = 9)
ggsave(file.path(base, "rc_curve_rejector_compare.svg"), p_rc, height = 5, width = 9)
ggsave(file.path(base, "marce_bar_rejector_compare.svg"), p_marce, height = 5, width = 9)
ggsave(file.path(base, "probs.svg"), p_probs_faceted, height = 6, width = 10)
if (exists("p_final_rr")) {
  ggsave(file.path(base, "rr_curve_final_deploy.svg"), p_final_rr, height = 5, width = 9)
}
if (exists("p_final_rc")) {
  ggsave(file.path(base, "rc_curve_final_deploy.svg"), p_final_rc, height = 5, width = 9)
}
```
