---
title: "MLL Lab validation cohort"
output: html_document
date: "`r format(Sys.Date())`"
---

Run first from repo root:

```bash
source .venv/bin/activate
python python/predict_new_samples.py \
  --input_file data/MLL_lab/STAR_AML_MLLlab.csv \
  --output_dir data/out/predictions \
  --cutoff_source both \
  --threshold_method both \
  --rejector_mode all \
  --max_accepted_risk_pct 5
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(purrr)
  library(knitr)
  library(kableExtra)
  library(caret)
})
```

```{r paths_helpers}
repo_root <- if (file.exists("R/utility_functions.R")) "." else ".."
source(file.path(repo_root, "R/mll_label_standardization.R"))
mapping <- MLL_ICC_MAPPING

mll_input_stem <- "STAR_AML_MLLlab"
mll_metadata_path <- file.path(repo_root, "data/MLL_lab/20231017_SampleMetadata.csv")
mll_pred_dir <- file.path(
  repo_root,
  "data/out/predictions",
  paste0(mll_input_stem, "_predictions")
)

# Filename suffix from predict_new_samples.py:
# {pred_suffix}_{rejector_key}[_deploy_loso][_pooled_oof][_risk5p0]

MLL_THRESHOLD_METHODS <- tibble::tribble(
  ~threshold_method,     ~threshold_tag,   ~threshold_label,
  "jackknife_adjusted",  "",               "Jackknife-adjusted",
  "pooled_oof",          "_pooled_oof",    "Pooled OOF",
  "ucb_95",              "_ucb_95",        "UCB (95%, one-sided)"
)

MLL_LABEL_SETS <- tibble::tribble(
  ~label_set_key,     ~pred_suffix,     ~label_display,    ~target_risk_pct, ~risk_tag,
  "merged_summed",    "merged_summed",  "Collapsed",       5.0,              "_risk5p0",
  "merged_maxprob",   "merged_maxprob", "Collapsed (max)", 5.0,              "_risk5p0",
  "unmerged_maxprob", "unmerged",       "Full subtypes",   5.0,              "_risk5p0"
)

MLL_PRED_CONFIGS <- tidyr::crossing(
  MLL_LABEL_SETS,
  tibble::tribble(
    ~rejector_key,                              ~rejector_label,
    "svm_single_head",                          "SVM maxprob single-head",
    "svm_ridge_in_model",                       "SVM ridge in-model"
  ),
  tibble::tribble(
    ~cutoff_source,     ~calibration_label, ~cutoff_tag,
    "selection_loso",   "Selection LOSO",   "",
    "deploy_loso",      "Deploy LOSO",      "_deploy_loso"
  ),
  MLL_THRESHOLD_METHODS
) %>%
  mutate(
    pred_filename = paste0(
      mll_input_stem,
      "_SVM_predictions_",
      pred_suffix, "_", rejector_key, cutoff_tag, threshold_tag, risk_tag,
      ".csv"
    ),
    pred_path = file.path(mll_pred_dir, pred_filename)
  )

build_mll_label_vocabulary <- function(label_set_key, label_display) {
  ref_path <- resolve_mll_pred_path(
    label_set_key, "svm_single_head", "selection_loso"
  )
  pred_df <- read.csv(ref_path, stringsAsFactors = FALSE)

  truth_vocab <- mll_meta %>%
    mutate(
      std_label = standardize_mll_truth(truth_raw, label_set_key),
      display_label = map_mll_to_display(std_label)
    ) %>%
    count(truth_raw, std_label, display_label, name = "n_truth")

  pred_vocab <- pred_df %>%
    mutate(
      pred_raw = as.character(prediction),
      std_label = standardize_mll_prediction(prediction, label_set_key),
      display_label = map_mll_to_display(std_label)
    ) %>%
    count(pred_raw, std_label, display_label, name = "n_pred")

  std_union <- sort(unique(c(truth_vocab$std_label, pred_vocab$std_label)))
  summary_tbl <- tibble::tibble(std_label = std_union) %>%
    left_join(
      truth_vocab %>%
        group_by(std_label, display_label) %>%
        summarise(n_truth = sum(n_truth), .groups = "drop"),
      by = "std_label"
    ) %>%
    left_join(
      pred_vocab %>%
        group_by(std_label, display_label) %>%
        summarise(n_pred = sum(n_pred), .groups = "drop") %>%
        select(std_label, n_pred),
      by = "std_label"
    ) %>%
    mutate(
      label_set = label_display,
      n_truth = tidyr::replace_na(n_truth, 0L),
      n_pred = tidyr::replace_na(n_pred, 0L),
      in_truth = n_truth > 0L,
      in_pred = n_pred > 0L,
      vocabulary_status = dplyr::case_when(
        in_truth & in_pred ~ "Both",
        in_truth & !in_pred ~ "Truth only",
        !in_truth & in_pred ~ "Prediction only",
        TRUE ~ "Neither"
      ),
      display_label = dplyr::coalesce(display_label, map_mll_to_display(std_label))
    ) %>%
    arrange(
      match(vocabulary_status, c("Truth only", "Prediction only", "Both")),
      desc(n_truth + n_pred),
      std_label
    )

  list(
    summary = summary_tbl,
    truth_detail = truth_vocab %>% mutate(label_set = label_display),
    pred_detail = pred_vocab %>% mutate(label_set = label_display)
  )
}

as_logical_flag <- function(x) {
  if (is.logical(x)) return(x)
  tolower(trimws(as.character(x))) %in% c("true", "t", "1", "yes", "y")
}

compute_kappa_safe <- function(truth, pred) {
  keep <- !is.na(truth) & !is.na(pred)
  truth <- as.character(truth[keep])
  pred <- as.character(pred[keep])
  if (length(truth) < 2L) return(NA_real_)
  lv <- sort(unique(c(truth, pred)))
  if (length(lv) < 2L) return(NA_real_)
  cm <- confusionMatrix(
    factor(pred, levels = lv),
    factor(truth, levels = lv)
  )
  as.numeric(cm$overall["Kappa"])
}

resolve_mll_pred_path <- function(
    cfg_label_set_key,
    cfg_rejector_key,
    cfg_cutoff_source = "selection_loso",
    cfg_threshold_method = "jackknife_adjusted",
    cfg_target_risk_pct = 5.0) {
  row <- MLL_PRED_CONFIGS %>%
    filter(
      .data$label_set_key == cfg_label_set_key,
      .data$rejector_key == cfg_rejector_key,
      .data$cutoff_source == cfg_cutoff_source,
      .data$threshold_method == cfg_threshold_method,
      .data$target_risk_pct == cfg_target_risk_pct
    )
  if (nrow(row) != 1L) {
    stop(sprintf(
      "Unknown MLL prediction config: %s / %s / %s / %s (target=%s)",
      cfg_label_set_key, cfg_rejector_key, cfg_cutoff_source,
      cfg_threshold_method, cfg_target_risk_pct
    ))
  }
  row$pred_path[[1]]
}
```

# Validation cohort analysis (MLL Lab)

## Load metadata and prediction outputs

```{r load_data}
if (!file.exists(mll_metadata_path)) {
  stop("Missing MLL metadata: ", mll_metadata_path)
}
if (!dir.exists(mll_pred_dir)) {
  stop(
    "Missing prediction directory: ", mll_pred_dir,
    ". Run predict_new_samples.py with --cutoff_source both --threshold_method both ",
    "--rejector_mode all --max_accepted_risk_pct 5."
  )
}

missing_preds <- MLL_PRED_CONFIGS %>%
  filter(!file.exists(.data$pred_path)) %>%
  select(pred_filename, pred_path)
if (nrow(missing_preds) > 0L) {
  stop(
    "Missing prediction file(s). Re-run predict_new_samples.py:\n",
    paste0("  - ", missing_preds$pred_path, collapse = "\n")
  )
}

MLL_lab <- read.csv(mll_metadata_path, sep = ";", stringsAsFactors = FALSE)
mll_meta <- MLL_lab %>%
  transmute(
    sample_name = as.character(MLL_ID),
    truth_raw = ICC_2022
  ) %>%
  filter(!is.na(truth_raw), nzchar(truth_raw))

# Classes absent from training (exclude from seen-coverage denominator).
mll_unseen_classes <- c("AML..NOS")

read_mll_prediction_metrics <- function(
    cfg_label_set_key,
    cfg_rejector_key,
    cfg_cutoff_source,
    cfg_threshold_method,
    cfg_target_risk_pct) {
  cfg <- MLL_PRED_CONFIGS %>%
    filter(
      .data$label_set_key == cfg_label_set_key,
      .data$rejector_key == cfg_rejector_key,
      .data$cutoff_source == cfg_cutoff_source,
      .data$threshold_method == cfg_threshold_method,
      .data$target_risk_pct == cfg_target_risk_pct
    )
  pred_df <- read.csv(cfg$pred_path[[1]], stringsAsFactors = FALSE) %>%
    mutate(
      sample_name = as.character(sample_name),
      pred_label = standardize_mll_prediction(prediction, cfg_label_set_key),
      passed_cutoff = as_logical_flag(prediction_passed_cutoff)
    ) %>%
    left_join(mll_meta, by = "sample_name") %>%
    mutate(truth_label = standardize_mll_truth(truth_raw, cfg_label_set_key)) %>%
    filter(!is.na(truth_label), !is.na(pred_label))

  if (nrow(pred_df) == 0L) return(NULL)

  pred_acc <- pred_df %>% filter(passed_cutoff)
  seen_mask <- !(pred_df$truth_label %in% mll_unseen_classes)

  tibble::tibble(
    threshold_method = cfg_threshold_method,
    threshold_label = cfg$threshold_label[[1]],
    label_set_key = cfg_label_set_key,
    label_display = cfg$label_display[[1]],
    rejector_key = cfg_rejector_key,
    rejector_label = cfg$rejector_label[[1]],
    cutoff_source = cfg_cutoff_source[[1]],
    calibration_label = cfg$calibration_label[[1]],
    target_risk_pct = cfg$target_risk_pct[[1]],
    pred_path = cfg$pred_path[[1]],
    n_samples = nrow(pred_df),
    n_accepted = nrow(pred_acc),
    coverage_pct = 100 * mean(pred_df$passed_cutoff, na.rm = TRUE),
    seen_coverage_pct = if (any(seen_mask, na.rm = TRUE)) {
      100 * mean(pred_df$passed_cutoff[seen_mask], na.rm = TRUE)
    } else {
      NA_real_
    },
    accepted_error_rate_pct = if (nrow(pred_acc) > 0L) {
      100 * mean(pred_acc$truth_label != pred_acc$pred_label, na.rm = TRUE)
    } else {
      NA_real_
    },
    kappa_without_rejection = compute_kappa_safe(pred_df$truth_label, pred_df$pred_label),
    kappa_with_rejection = compute_kappa_safe(pred_acc$truth_label, pred_acc$pred_label)
  )
}

mll_metrics <- purrr::pmap_dfr(
  MLL_PRED_CONFIGS %>%
    select(
      label_set_key, rejector_key, cutoff_source,
      threshold_method, target_risk_pct
    ),
  function(label_set_key, rejector_key, cutoff_source, threshold_method, target_risk_pct) {
    read_mll_prediction_metrics(
      label_set_key, rejector_key, cutoff_source,
      threshold_method, target_risk_pct
    )
  }
)

knitr::kable(
  mll_metrics %>%
    select(
      threshold_label, calibration_label, label_display,
      rejector_label, target_risk_pct, n_samples, n_accepted
    ) %>%
    mutate(
      label_display = as.character(label_display),
      target_risk_pct = sprintf("%.1f%%", target_risk_pct)
    ),
  caption = "Loaded MLL prediction runs (threshold x calibration track x rejector)"
)
```

## Label vocabulary: MLL truth vs model prediction classes

Standardized labels use the same rules as training (`merge_true_labels` for collapsed classes; full ICC names for unmerged). KMT2A::MLLT3 fusion order is harmonized to the training canonical string. Model rare-fusion classes (e.g. `CBFA2T3..GLIS2`, `KAT6A..CREBBP`) are bucketed to `AML.other.rare.transloc` to match MLL ICC truth. Non-MLLT3 KMT2A model fusions (e.g. `KMT2A..ELL`) are bucketed to Other KMT2A (`AML.otherKMT2A` unmerged; `other.KMT2A` collapsed).

```{r label_vocabulary}
mll_label_vocab_collapsed <- build_mll_label_vocabulary("merged_summed", "Collapsed")
mll_label_vocab_unmerged <- build_mll_label_vocabulary("unmerged_maxprob", "Full subtypes")

mll_label_vocab_summary <- bind_rows(
  mll_label_vocab_collapsed$summary,
  mll_label_vocab_unmerged$summary
)

mll_label_vocab_mismatches <- mll_label_vocab_summary %>%
  filter(.data$vocabulary_status != "Both")

knitr::kable(
  mll_label_vocab_summary %>%
    select(
      label_set, vocabulary_status, display_label, std_label, n_truth, n_pred
    ),
  caption = "Standardized label vocabulary on MLL (truth metadata vs primary model predictions)"
)

if (nrow(mll_label_vocab_mismatches) > 0L) {
  knitr::kable(
    mll_label_vocab_mismatches %>%
      select(label_set, vocabulary_status, display_label, std_label, n_truth, n_pred),
    caption = "Labels present in only truth or only predictions (potential naming / OOD gaps)"
  )
}

knitr::kable(
  mll_label_vocab_collapsed$truth_detail %>%
    select(truth_raw, std_label, display_label, n_truth) %>%
    arrange(desc(n_truth)),
  caption = "Collapsed label set: ICC truth codes mapped to model label space"
)

knitr::kable(
  mll_label_vocab_collapsed$pred_detail %>%
    select(pred_raw, std_label, display_label, n_pred) %>%
    arrange(desc(n_pred)),
  caption = "Collapsed label set: raw model prediction classes"
)

knitr::kable(
  mll_label_vocab_unmerged$truth_detail %>%
    select(truth_raw, std_label, display_label, n_truth) %>%
    arrange(desc(n_truth)),
  caption = "Full subtypes: ICC truth codes mapped to model label space"
)

knitr::kable(
  mll_label_vocab_unmerged$pred_detail %>%
    select(pred_raw, std_label, display_label, n_pred) %>%
    arrange(desc(n_pred)),
  caption = "Full subtypes: raw model prediction classes"
)

dir.create(file.path(repo_root, "writing/tables_new"), recursive = TRUE, showWarnings = FALSE)
write.csv(
  mll_label_vocab_summary,
  file.path(repo_root, "writing/tables_new/mll_label_vocabulary_summary.csv"),
  row.names = FALSE
)
write.csv(
  bind_rows(
    mll_label_vocab_collapsed$truth_detail %>% mutate(source = "truth"),
    mll_label_vocab_collapsed$pred_detail %>% mutate(source = "prediction")
  ),
  file.path(repo_root, "writing/tables_new/mll_label_vocabulary_collapsed_detail.csv"),
  row.names = FALSE
)
write.csv(
  bind_rows(
    mll_label_vocab_unmerged$truth_detail %>% mutate(source = "truth"),
    mll_label_vocab_unmerged$pred_detail %>% mutate(source = "prediction")
  ),
  file.path(repo_root, "writing/tables_new/mll_label_vocabulary_unmerged_detail.csv"),
  row.names = FALSE
)
```

## Table 2: MLL lab cohort distribution

```{r table2, results='asis'}
MLL_lab_df <- mll_meta %>%
  transmute(
    Subtype = map_truth_to_canonical(truth_raw, mapping),
    Cohort = "MLL Munich"
  )

MLL_lab_counts <- as.data.frame(table(MLL_lab_df$Subtype))
colnames(MLL_lab_counts) <- c("Subtype", "MLL Munich")
MLL_lab_counts <- MLL_lab_counts[order(-MLL_lab_counts$`MLL Munich`), ]

MLL_lab_counts$Subtype <- as.character(MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("\\.", " ", MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("AML with inv 16  t 16 16  CBFB  MYH11", "CBFB::MYH11", MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("APL  t 15 17  PML  RARA", "PML::RARA", MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("AML with t 6 9  DEK  NUP214", "DEK::NUP214", MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("AML with mutated NPM1", "Mutated NPM1", MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("AML with t 8 21  RUNX1  RUNX1T1", "RUNX1::RUNX1T1", MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("AML with in frame bZIP CEBPA", "CEBPA bZIP in-frame", MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("AML with t 9 11  MLLT3  KMT2A", "KMT2A::MLLT3", MLL_lab_counts$Subtype)
MLL_lab_counts$Subtype <- gsub("MDS", "MDS-related", MLL_lab_counts$Subtype)

total_row <- data.frame(
  Subtype = "Total",
  `MLL Munich` = sum(MLL_lab_counts$`MLL Munich`),
  stringsAsFactors = FALSE,
  check.names = FALSE
)
MLL_lab_counts_with_total <- rbind(MLL_lab_counts, total_row)

mll_table <- kable(
  MLL_lab_counts_with_total,
  format = "html",
  caption = "Table 2: Distribution of leukemia subtypes in MLL lab cohort",
  row.names = FALSE,
  escape = FALSE
) %>%
  kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(nrow(MLL_lab_counts_with_total), bold = TRUE, background = "#e6f2ff")

mll_table
dir.create(file.path(repo_root, "writing/tables_new"), recursive = TRUE, showWarnings = FALSE)
save_kable(mll_table, file.path(repo_root, "writing/tables_new/table2.png"), zoom = 2)
```

## MLL performance by calibration track, label set, and rejector

Deployed cutoffs at **5%** target risk (collapsed and full subtypes). Metrics on samples with ICC truth labels.

```{r performance_table}
mll_perf_table <- mll_metrics %>%
  mutate(
    target_risk_pct = sprintf("%.1f%%", target_risk_pct),
    accepted_error_rate_pct = round(accepted_error_rate_pct, 2),
    coverage_pct = round(coverage_pct, 2),
    seen_coverage_pct = round(seen_coverage_pct, 2),
    kappa_without_rejection = round(kappa_without_rejection, 3),
    kappa_with_rejection = round(kappa_with_rejection, 3)
  ) %>%
  select(
    threshold_label,
    calibration_label,
    label_display,
    rejector_label,
    target_risk_pct,
    accepted_error_rate_pct,
    coverage_pct,
    seen_coverage_pct,
    kappa_without_rejection,
    kappa_with_rejection,
    n_samples,
    n_accepted
  ) %>%
  arrange(threshold_label, label_display, rejector_label, calibration_label)

knitr::kable(
  mll_perf_table,
  align = c("l", "l", "l", "l", "r", "r", "r", "r", "r", "r", "r", "r"),
  col.names = c(
    "Cutoff method",
    "Calibration",
    "Label set",
    "Rejector",
    "Target risk",
    "Accepted error (%)",
    "Coverage (%)",
    "Seen coverage (%)",
    "Kappa (all)",
    "Kappa (accepted)",
    "N",
    "N accepted"
  ),
  caption = "MLL lab performance: SVM classifier + SVM rejectors at 5% target risk"
)

dir.create(file.path(repo_root, "writing/tables_new"), recursive = TRUE, showWarnings = FALSE)
write.csv(
  mll_perf_table,
  file.path(repo_root, "writing/tables_new/mll_perf_table.csv"),
  row.names = FALSE
)
write.csv(
  mll_perf_table,
  file.path(repo_root, "writing/tables_new/mll_calibration_track_svm_comparison.csv"),
  row.names = FALSE
)
```

## Side-by-side delta (deploy-loso minus selection-loso)

```{r calibration_delta}
mll_calibration_delta <- mll_metrics %>%
  select(
    threshold_method, threshold_label,
    label_set_key, label_display, target_risk_pct,
    rejector_key, rejector_label, cutoff_source,
    accepted_error_rate_pct, coverage_pct, seen_coverage_pct,
    kappa_without_rejection, kappa_with_rejection, n_accepted
  ) %>%
  tidyr::pivot_wider(
    names_from = cutoff_source,
    values_from = c(
      accepted_error_rate_pct, coverage_pct, seen_coverage_pct,
      kappa_without_rejection, kappa_with_rejection, n_accepted
    )
  ) %>%
  mutate(
    delta_accepted_error_pct = accepted_error_rate_pct_deploy_loso - accepted_error_rate_pct_selection_loso,
    delta_coverage_pct = coverage_pct_deploy_loso - coverage_pct_selection_loso,
    delta_seen_coverage_pct = seen_coverage_pct_deploy_loso - seen_coverage_pct_selection_loso,
    delta_kappa_accepted = kappa_with_rejection_deploy_loso - kappa_with_rejection_selection_loso,
    delta_n_accepted = n_accepted_deploy_loso - n_accepted_selection_loso
  ) %>%
  select(
    threshold_label, label_display, target_risk_pct, rejector_label,
    accepted_error_rate_pct_selection_loso, accepted_error_rate_pct_deploy_loso, delta_accepted_error_pct,
    coverage_pct_selection_loso, coverage_pct_deploy_loso, delta_coverage_pct,
    kappa_with_rejection_selection_loso, kappa_with_rejection_deploy_loso, delta_kappa_accepted,
    n_accepted_selection_loso, n_accepted_deploy_loso, delta_n_accepted
  ) %>%
  mutate(target_risk_pct = sprintf("%.1f%%", target_risk_pct)) %>%
  arrange(threshold_label, label_display, rejector_label)

knitr::kable(
  mll_calibration_delta %>%
    mutate(across(where(is.numeric), ~round(.x, 3))),
  caption = "Deploy-loso vs selection-loso on MLL (positive delta = deploy-loso higher; per threshold method)"
)

write.csv(
  mll_calibration_delta,
  file.path(repo_root, "writing/tables_new/mll_calibration_track_svm_delta.csv"),
  row.names = FALSE
)
```

## Side-by-side delta (pooled OOF minus jackknife-adjusted)

Selection-loso cutoffs; same rejector.

```{r threshold_delta}
mll_threshold_delta <- mll_metrics %>%
  filter(.data$cutoff_source == "selection_loso") %>%
  select(
    label_display, target_risk_pct, rejector_label, threshold_method,
    accepted_error_rate_pct, coverage_pct, seen_coverage_pct,
    kappa_with_rejection, n_accepted
  ) %>%
  tidyr::pivot_wider(
    names_from = threshold_method,
    values_from = c(
      accepted_error_rate_pct, coverage_pct, seen_coverage_pct,
      kappa_with_rejection, n_accepted
    )
  ) %>%
  mutate(
    delta_accepted_error_pct = accepted_error_rate_pct_pooled_oof - accepted_error_rate_pct_jackknife_adjusted,
    delta_coverage_pct = coverage_pct_pooled_oof - coverage_pct_jackknife_adjusted,
    delta_kappa_accepted = kappa_with_rejection_pooled_oof - kappa_with_rejection_jackknife_adjusted,
    delta_n_accepted = n_accepted_pooled_oof - n_accepted_jackknife_adjusted
  ) %>%
  select(
    label_display, target_risk_pct, rejector_label,
    accepted_error_rate_pct_jackknife_adjusted, accepted_error_rate_pct_pooled_oof, delta_accepted_error_pct,
    coverage_pct_jackknife_adjusted, coverage_pct_pooled_oof, delta_coverage_pct,
    kappa_with_rejection_jackknife_adjusted, kappa_with_rejection_pooled_oof, delta_kappa_accepted,
    n_accepted_jackknife_adjusted, n_accepted_pooled_oof, delta_n_accepted
  ) %>%
  mutate(target_risk_pct = sprintf("%.1f%%", target_risk_pct)) %>%
  arrange(label_display, rejector_label)

knitr::kable(
  mll_threshold_delta %>%
    mutate(across(where(is.numeric), ~round(.x, 3))),
  caption = "Pooled-OOF minus jackknife-adjusted cutoffs on MLL (positive delta = pooled OOF higher)"
)

write.csv(
  mll_threshold_delta,
  file.path(repo_root, "writing/tables_new/mll_threshold_method_delta.csv"),
  row.names = FALSE
)
```

## Confusion matrices (collapsed, both rejectors)

```#{r confusion_collapsed, fig.width=12, fig.height=5}
load_mll_pred_joined <- function(
    label_set_key,
    rejector_key,
    cutoff_source = "selection_loso",
    threshold_method = "jackknife_adjusted") {
  pred_path <- resolve_mll_pred_path(
    label_set_key, rejector_key, cutoff_source, threshold_method
  )
  read.csv(pred_path, stringsAsFactors = FALSE) %>%
    mutate(
      sample_name = as.character(sample_name),
      pred_label = standardize_mll_prediction(prediction, label_set_key),
      passed_cutoff = as_logical_flag(prediction_passed_cutoff)
    ) %>%
    left_join(mll_meta, by = "sample_name") %>%
    mutate(
      truth_label = standardize_mll_truth(truth_raw, label_set_key),
      truth_display = map_mll_to_display(truth_label),
      pred_display = map_mll_to_display(pred_label)
    ) %>%
    filter(!is.na(truth_label), !is.na(pred_label))
}

plot_confusion_heatmap <- function(df, title) {
  if (nrow(df) == 0L) {
    return(
      ggplot2::ggplot() +
        ggplot2::annotate(
          "text", x = 0.5, y = 0.5,
          label = "No accepted samples",
          size = 4
        ) +
        ggplot2::xlim(0, 1) +
        ggplot2::ylim(0, 1) +
        ggplot2::theme_void() +
        ggplot2::labs(title = title)
    )
  }

  cm_df <- as.data.frame(
    table(
      Truth = df$truth_display,
      Prediction = df$pred_display,
      useNA = "no"
    ),
    stringsAsFactors = FALSE
  )
  names(cm_df) <- c("Truth", "Prediction", "Freq")
  cm_df <- cm_df %>% filter(.data$Freq > 0L)

  ggplot2::ggplot(cm_df, ggplot2::aes(x = .data$Prediction, y = .data$Truth, fill = .data$Freq)) +
    ggplot2::geom_tile(color = "white") +
    ggplot2::geom_text(ggplot2::aes(label = .data$Freq), color = "black", size = 3) +
    ggplot2::scale_fill_gradient(low = "#f7fbff", high = "#08306b") +
    ggplot2::theme_bw() +
    ggplot2::theme(
      axis.text.x = ggplot2::element_text(angle = 45, hjust = 1),
      legend.position = "right"
    ) +
    ggplot2::labs(title = title, x = NULL, y = NULL, fill = "N")
}

suppressPackageStartupMessages(library(ggplot2))
library(patchwork)

cm_plots <- lapply(seq_len(nrow(mll_metrics)), function(i) {
  row <- mll_metrics[i, ]
  df <- load_mll_pred_joined(
    row$label_set_key, row$rejector_key, row$cutoff_source, row$threshold_method
  )
  df_acc <- df %>% filter(passed_cutoff)
  plot_confusion_heatmap(
    df_acc,
    sprintf(
      "%s | %s | %s | %s (accepted only, n=%d)",
      row$calibration_label, row$threshold_label, row$label_display, row$rejector_label, nrow(df_acc)
    )
  )
})

wrap_plots(cm_plots, ncol = 2)
```

## Sample-level export (primary collapsed run)

Primary run: collapsed labels, SVM maxprob single-head, selection-loso cutoffs, jackknife-adjusted threshold.

```{r export_primary}
primary_path <- resolve_mll_pred_path(
  "merged_summed", "svm_single_head", "selection_loso"
)
MLL_lab_pred_df_merged <- read.csv(primary_path, stringsAsFactors = FALSE)

MLL_lab_sample_df <- load_mll_pred_joined(
  "merged_summed", "svm_single_head", "selection_loso"
) %>%
  transmute(
    sample_name,
    truth_display,
    prediction_display = pred_display,
    passed_cutoff,
    prediction_prob,
    prediction_prob_calibrated,
    correct = truth_display == pred_display
  )

write.csv(
  MLL_lab_sample_df,
  file.path(repo_root, "data/out/predictions/STAR_AML_MLLlab_predictions/mll_sample_summary_collapsed_svm_single_head.csv"),
  row.names = FALSE
)

knitr::kable(
  head(MLL_lab_sample_df, 10),
  caption = "First 10 samples (collapsed label set, SVM maxprob single-head rejector)"
)
```
